---
title: "lIPIDpeROXIDATION.Rmd"
author: "Sam Gurr"
date: "2024-10-31"
output: html_document
---

-   Last updates: 11 13, 2024

# Thermal Performacne (pilot test) & INFISH projet with Indya Limpkin & Ashley Hallengren

## Ran a Lipid peroxidation kit on C virginica gill tissue homogenate, corrected to

total protein using pierce BCA protein assay (albumin standard)

### SET UP

* note: add your own directory below to navigate to RAnalysis folder containing the data, output, and script subfolders 
that reflect the google drive. alterntively clone the repository EAD-ASEB-Cvirginica_Thermal_Performance from github
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
# knitr::opts_knit$set(root.dir = "C:/Users/katherine.mcfarland/Documents/GitHub/EAD-ASEB-Cvirginica_Thermal_Performance/larvae") # Katie's
knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Cvirginica_Thermal_Performance/RAnalysis") # Sam's
#knitr::opts_knit$set(root.dir = "C:/Users/samuel.gurr/Documents/Github_repositories/EAD-ASEB-Cvirginica_Thermal_Performance/RAnalysis") # Sam's work

```


## Load packages 

* use install.packges('pacakge name') if you do not have these downloaded
```{r setup, include=TRUE}

library(ggpubr)
library(ggplot2)
library(dplyr)
library(lmtest) # to receive p value from betareg model
# library(FSA) # for the Dun test post hoc for SRH non-parametric 2 way anova]
library(emmeans)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(car)
library(lmerTest)
library(tidyr)
library(reshape2)
library(ggpubr)
library(nlme)
library(rcompanion) # to run the Schrier -Ray-Hare non parametric 2 way 
library(ggpmisc) # stat_poly for inserting equation and R2 for ggplot line 
```

## Function to load files

-   At your service, from 'spec to table' ...I'm so funny

```{r Spec_to_table}

Spec_to_table <- function(datapath, platenumber, daterecorded) {
  
# import the data in raw text form exported from the spectrophotometer
        rawtextfile <- read.delim(datapath,
                              header = F,
                              skip=3,
                              fileEncoding="latin1")
  

        rawmatrix   <- as.matrix(rawtextfile[c(1:8),c(3:14)])
        colnames(rawmatrix) = c("1","2","3","4","5","6","7","8","9","10","11","12")
        rownames(rawmatrix) = c("A","B","C","D","E","F","G","H")
        
        raw_table <- as.data.frame.table(rawmatrix, responseName = "value") %>% 
                                    dplyr::rename(well_row=Var1, well_column=Var2, Abs_412nm=value) %>% 
                                    dplyr::mutate(well=paste0(well_row,well_column),
                                                  Run_date =daterecorded,
                                                  Plate=platenumber) %>% 
                                    dplyr::select(-c(well_row,well_column))
        # reorder the columns because I prefer it this way..
        
        raw_table_ordered <- raw_table[,c(3,4,2,1)]
}


```

-   Use the spec_to_table to import raw spec txt files and compile to a table, yes that is what it does

-   add an "about" column - we have a few hiccups with this assay and reran it with dilutions

  - no dilution = ran without additional dilutions 
  - 5x dilution post = after we ran the no dilution we saw any values were ut of range (too high) we diluted 5x with ultrapure water direct to the plate and reran
  - 5x dilution post 30 mins = same as the 5 x dilution but after a 30 minute incubation we ran it again

```{r load Succinate Synthase raw data and build master files}

# plate 1 
MDA_Plate1 <- Spec_to_table('Data/Lipid_Peroxidation/20241016_LipidPeroxidation_Plate1.txt', 1, 20241016)

# plate 2
MDA_Plate2 <- Spec_to_table('Data/Lipid_Peroxidation/20241016_LipidPeroxidation_Plate2.txt', 2, 20241016)

```

```{r load Total protein correction data for succinate}

# plate 1 
TP_Plate1_master <- Spec_to_table('Data/Total_Protein_Lipid_Peroxidation/20241017_TotalProtein_LipidPeroxidation_Plate1_37C.txt', 1, 20241017)  %>%  dplyr::rename(Abs_562nm = Abs_412nm)

# plate 2
TP_Plate2_master <- Spec_to_table('Data/Total_Protein_Lipid_Peroxidation/20241017_TotalProtein_LipidPeroxidation_Plate2_37C.txt', 2, 20241017)   %>%  dplyr::rename(Abs_562nm = Abs_412nm)

```


```{r load reference metadata and call target datasets}

metadata <- read.csv('Data/Lipid_peroxidation_metadata.csv', sep = ",", header=T)

# rename a few columns so they can merge with the data (in next chunk)

# Plate 1 
metadata.MDA_Plate_1 <- metadata %>% 
                        dplyr::filter(Assay.Type %in% "Lipid Peroxidation" & Plate.Number == 1) %>% 
                        dplyr::rename(Plate = Plate.Number,
                                      well  = Well.ID)
# Plate 2
metadata.MDA_Plate_2 <- metadata %>% 
                        dplyr::filter(Assay.Type %in% "Lipid Peroxidation" & Plate.Number == 2) %>% 
                        dplyr::rename(Plate = Plate.Number,
                                      well  = Well.ID)



# Plate 1 
metadata.TP_Plate_1 <- metadata %>% 
                        dplyr::filter(Assay.Type %in% "Total Protein" & Plate.Number == 1) %>% 
                        dplyr::rename(Plate = Plate.Number,
                                      well  = Well.ID)
# Plate 2
metadata.TP_Plate_2 <- metadata %>% 
                        dplyr::filter(Assay.Type %in% "Total Protein" & Plate.Number == 2) %>% 
                        dplyr::rename(Plate = Plate.Number,
                                      well  = Well.ID)



```


* merge metadata with data 
```{r merge metadata with data}

# MDA
# Plate 1 
Plate1 <- merge(metadata.MDA_Plate_1, MDA_Plate1, by = c('Plate', 'well'))                          
nrow(MDA_Plate1) == nrow(Plate1) # must be TRUE
Plate1 <- subset(Plate1, !is.na(Sample.Type)) # 96 rows


# Plate 2
Plate2 <- merge(metadata.MDA_Plate_2, MDA_Plate2, by = c('Plate', 'well'))                          
nrow(MDA_Plate2) == nrow(Plate2) # must be TRUE - NOT HERE THOUGH because metadtata did not include the empty IDs, we have 78 with sample, all good
Plate2 <- subset(Plate2, !is.na(Sample.Type)) # now 68 with NAs excluded


# Total protein
# Plate 1 
Plate1.TP <- merge(metadata.TP_Plate_1, TP_Plate1_master, by = c('Plate', 'well'))                          
nrow(TP_Plate1_master) == nrow(Plate1.TP) # must be TRUE
Plate1.TP <- subset(Plate1.TP, !is.na(Sample.Type)) # 96 rows

# Plate 2 
Plate2.TP <- merge(metadata.TP_Plate_2, TP_Plate2_master, by = c('Plate', 'well'))                          
nrow(TP_Plate2_master) == nrow(Plate2.TP) # must be TRUE - NOT HERE THOUGH 
Plate2.TP <- subset(Plate2.TP, !is.na(Sample.Type)) # now 80 total rows with NAs omitted (not sample, blank nor stnadard)


```


* subset standards, blanks, and samples
```{r subset data types}

# Citrate synthase 
# Plate 1 
Plate1.samples   <- Plate1 %>% dplyr::filter(Sample.Type %in% 'Sample') %>% dplyr::rename(Abs_532nm = Abs_412nm)
# Plate1.blanks    <- Plate1 %>% dplyr::filter(Sample.Type %in% 'Blank') # space, why... whatevs
Plate1.standards <- Plate1 %>% dplyr::filter(Sample.Type %in% c('Blank','Standard')) %>% 
                               dplyr::rename(Abs_532nm = Abs_412nm) %>% 
                               dplyr::mutate(nmolMDA = # ashely input metadata for the u; standard, convert to the uM MDA
                                              case_when(Standards == 0 ~ 0, 
                                                        Standards == 6 ~ 2,
                                                        Standards == 12 ~ 4, 
                                                        Standards == 18 ~ 6, 
                                                        Standards == 24 ~ 8, 
                                                        Standards == 30 ~ 10))

(nrow(Plate1.samples) + 
    # nrow(Plate1.blanks) + 
      nrow(Plate1.standards)) == nrow(Plate1) # sanity check, must be true


# Plate 2
Plate2.samples   <- Plate2 %>% dplyr::filter(Sample.Type %in% 'Sample') %>% dplyr::rename(Abs_532nm = Abs_412nm)
# Plate2.blanks    <- Plate2 %>% dplyr::filter(Sample.Type %in% 'Blank')  # space, why... whatevs
Plate2.standards <- Plate2 %>% dplyr::filter(Sample.Type %in% c('Blank','Standard')) %>% 
                               dplyr::rename(Abs_532nm = Abs_412nm) %>% 
                               dplyr::mutate(nmolMDA = # ashely input metadata for the u; standard, convert to the uM MDA
                                              case_when(Standards == 0 ~ 0, 
                                                        Standards == 6 ~ 2,
                                                        Standards == 12 ~ 4, 
                                                        Standards == 18 ~ 6, 
                                                        Standards == 24 ~ 8, 
                                                        Standards == 30 ~ 10))
                                

(nrow(Plate2.samples) + 
    # nrow(Plate2.blanks) + 
    nrow(Plate2.standards)) == nrow(Plate2) # sanity check, must be true






# Total protein
# Plate 1 
Plate1_TP.samples    <- Plate1.TP %>% dplyr::filter(Sample.Type %in% 'Sample')
Plate1_TP.blanks     <- Plate1.TP %>% dplyr::filter(Sample.Type %in% 'Blank')
Plate1_TP.standards  <- Plate1.TP %>% dplyr::filter(Sample.Type %in% c('Standard', 'Blank'))

(nrow(Plate1_TP.samples) + 
    # nrow(Plate1_TP.blanks) + 
      nrow(Plate1_TP.standards)) == nrow(Plate1.TP) # sanity check, must be true


# Plate 2
Plate2_TP.samples    <- Plate2.TP %>% dplyr::filter(Sample.Type %in% 'Sample')
Plate2_TP.blanks     <- Plate2.TP %>% dplyr::filter(Sample.Type %in% 'Blank') 
Plate2_TP.standards  <- Plate2.TP %>% dplyr::filter(Sample.Type %in% c('Standard', 'Blank'))

(nrow(Plate2_TP.samples) + 
    # nrow(Plate2_TP.blanks) + 
      nrow(Plate2_TP.standards)) == nrow(Plate2.TP) # sanity check, must be true

```

## Total protein calculations 

```{r total protein standard curve}

# assign known BCA values to standard IDs A-I
BCA_standards <- rbind(Plate1_TP.standards, Plate2_TP.standards) %>%  
  dplyr::mutate(BCA_ug_mL = case_when(Standards %in% 'A' ~ 2000,
                                      Standards %in% 'B' ~ 1500,
                                      Standards %in% 'C' ~ 1000,
                                      Standards %in% 'D' ~ 750,
                                      Standards %in% 'E' ~ 500,
                                      Standards %in% 'F' ~ 250,
                                      Standards %in% 'G' ~ 125,
                                      Standards %in% 'H' ~ 25,
                                      Standards %in% 'I' ~ 0)) %>% 
  dplyr::select(Plate, Standards, BCA_ug_mL, Abs_562nm)

# Run standard curve, calculate totalprotein 
BCA_background_zero <- BCA_standards %>% 
                        dplyr::filter(Standards %in% 'I') %>% # the zero standard
                        dplyr::group_by(Plate, Standards, BCA_ug_mL) %>% # group by to get the means
                        dplyr::summarise_each(funs(mean,sd,se=sd(.)/sqrt(n()))) # get all the stats 


# Plate 1, blank to correct by is 0.07925
# Plate 2, blank to correct by is 0.08185



# Absorbance corrected - take the mean of any duplicates
BCA_standards_means <- BCA_standards %>% 
                        dplyr::filter(!BCA_ug_mL %in% 0) %>% 
                        dplyr::mutate(Abs_562nm_cor = 
                                      case_when(Plate == 1 ~ (Abs_562nm-0.07925),
                                                Plate == 2 ~ (Abs_562nm-0.08185) ) ) %>% 
                        dplyr::select(-Abs_562nm) %>% 
                        dplyr::group_by(Plate, Standards, BCA_ug_mL) %>% # group by to get the means
                        dplyr::summarise_each(funs(mean,sd,se=sd(.)/sqrt(n()))) # get all the stats 


# plot it insert the quadratic formaula using ggpmisc
BCA_stand_plots_quadratic <- BCA_standards_means %>% 
                    # QUADRATIC SMOOTH LINE WORKS BEST HERE (MANUFACTURERS INSTRUCTIONS)
                     # dplyr::filter(!(Plate %in% 2 & Standards %in% 'D')) %>% # hash me out to test
                     ggplot(aes(y=mean, x=BCA_ug_mL)) + 
                        geom_point() +
                        theme_bw() +
                        labs(y= "Net Abs 562nm", x = "Protein Concentration in ug/mL") +
                        #geom_line() +
                        #stat_poly_line(color='red') +
                        #geom_smooth() +
                        stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1) +
                        stat_poly_eq(parse=T, aes(label = ..eq.label..), formula=y ~ x + I(x^2)) +
                        ggtitle('Total protein: Quadratic curve') +
                        #stat_poly_eq(use_label(c("eq", "R2"))) +
                        facet_wrap(~Plate) 

# output 
pdf(paste("Output/Colorimetric_assays/Lipid_Peroxidation/Standard_Curve_BCA_MDAcorrection.pdf", sep =''), 
    width=10, 
    height=7)
print(ggarrange(BCA_stand_plots_quadratic))
dev.off()

```


* Use BCA standard curves to calculate TP 

```{r calc TP}
library(grDevices)
# Standard curve, Plate 1 equation y = -0.0189 + 0.00101x - 7.71x10^-8x^2 - need to solve for x!
# Standard curve, Plate 2 equation y = --0.0189 + 0.00196x - 4.42x10^-7x^2 - need to solve for x!

# Standard curve, Plate 1
a1 <- -1.92*10^-7
b1 <- 0.00159
c1 <- -0.0158
# EQ: (-(b1) + sqrt( (b1^2) - (4*(((a1)-Abs_562nm_cor))*(c1)) ))/(2*a1)

# Standard curve, Plate 2
a2 <- -2*10^-7
b2 <- 0.00152
c2 <- 0.00093
# EQ: (-(b2) + sqrt( (b2^2) - (4*a2*(c2-Abs_562nm_cor)) ) ) / (2*a2)


# linear equation plate 1 == (Abs_562nm_cor - 0.192)/0.000993
# linear equation plate 2 == (Abs_562nm_cor - 0.224)/0.000911


# IMPORTANT! we used 25 ul of the standards and 25 ul of the unknowns (samples) 
# therefore we can interpret the unknown direct to the the standard curve without having 
# to account for addition factors, fot example, if we used 5 ul unknown (sample) we would have to adjust 
# by multiplying by 5 to reach the standard curve 

V = 0.025 # 25 ul or 0.025 mL

# Again, remember Plate 1, blank to correct by is 0.1372
# Again, remember Plate 2, blank to correct by is 0.1312

# Sanity check Lets look at the absorbance vs. totla protein concentration data 

TotalProtein_final <- rbind(Plate1_TP.samples, Plate2_TP.samples) %>% 
                      dplyr::filter(!Abs_562nm >= 4) %>% 
                      dplyr::select(Plate,Tank.Number,Temperature.Number,Abs_562nm) %>% 
                      # dplyr::filter(!Abs_562nm > 3.99) %>% # data as 4.00 is above the detection limit, omit
                      dplyr::mutate(Unique_ID = 
                                      paste0('Tank',Tank.Number, '_',
                                             'Temperature', Temperature.Number)) %>% # unique ID t0 group by
                      dplyr::mutate(Abs_562nm_cor = # correct the raw abs, subtract background
                                   case_when(Plate == 1 ~ (Abs_562nm-0.07925), # for plate 1
                                             Plate == 2 ~ (Abs_562nm-0.08185) ) ) %>% # for plate 2 
                      dplyr::mutate(TotalProtein_ug_mL = 
                                    case_when(
                                      # linear fr neg discrim. - luckily only two values from plate 2
                                      # Scallop_ID %in% c(33, 51) ~ 
                                      #   ((Abs_562nm_cor - 0.224)/0.000911),
                                      # quadratic for Plate 1
                                      Plate == 1 ~ 
                                        ((-(b1) + sqrt( (b1^2) - (4*a1*(c1-Abs_562nm_cor)) ) ) / (2*a1)), 
                                      # quadratic for plate 2
                                      Plate == 2 ~ 
                                        ((-(b2) + sqrt( (b2^2) - (4*a2*(c2-Abs_562nm_cor)) ) ) / (2*a2)) ),
                                    # ug per mL concentration to ug in 25 ul sample 
                                    TotalProtein_ug = TotalProtein_ug_mL*V) %>% 
                      na.omit()

length(unique(TotalProtein_final$Unique_ID)) # 36 samples retained for total protein

# View(TotalProtein_final)
nrow(TotalProtein_final) # 45

# NOTE! I found that some of my discriminants are negative (b^2 - 4ac) IN PLATE 2!! (as 'NaN)
# so Im going to extrapolate from a linear regression curve based on the final samples 

# first lets plot absorbance and actual concentration calacuated 
calc_BCA_plot <- TotalProtein_final %>% 
                    ggplot(aes(y = Abs_562nm_cor, 
                               x  = TotalProtein_ug_mL)) +
                          geom_point() + 
                    theme_bw() + 
                    ggtitle('Total protein: Calculated BCA by Net Absorbance') +
                    facet_wrap(~Plate)
calc_BCA_plot 

#print
pdf(paste("Output/Colorimetric_assays/Lipid_Peroxidation/Standard_Curve_BCA_MDAcorrection.pdf", sep =''), 
    width=10, 
    height=14)
ggpubr::ggarrange(BCA_stand_plots_quadratic,
          calc_BCA_plot,
          nrow = 2)
dev.off()
# write csv
write.csv(TotalProtein_final, file = "Output/Colorimetric_assays/Lipid_Peroxidation/TotalProteinBCA_MDAcorrection.csv")


```


###  THIS IS WHERE I LEFT OFF ON 11/13/2024 
* NOTE: the code below is from the succinate script, need to change throughout to accomodate the variables and values in chunks above


## Lipid Peroxidation Calculations


* STEP 1: Blanks

  - calc the mean blanks by time and plate. 
  
  - **note** this is not a normal blank, but the change of the blank 
  over time final - time initial. This is important downstream in the calculation, youll see later..

  -  **remember** we recorded every 5 minutes unitl the samples surpased the highest standard
we only need timepoint initial and timepoint end, lets average for each and truncate later..
```{r blank rate}

BCA_background_zero <- BCA_standards %>% 
                        dplyr::filter(Standards %in% 'I') %>% # the zero standard
                        dplyr::group_by(Plate, Standards, BCA_ug_mL) %>% # group by to get the means
                        dplyr::summarise_each(funs(mean,sd,se=sd(.)/sqrt(n()))) # get all the stats 

Plate1.blanks.means <- BCA_background_zero[1,]
Plate2.blanks.means <- BCA_background_zero[2,]
# Plate 1, blank to correct by is 0.07925
# Plate 2, blank to correct by is 0.08185

```



* STEP 2: Correct for background

  - From protocol (Fluorometric protocol was used!): Determine the slope of the standard florescence values 
  and calculate the the succinate concetration as follows

SUccinate (uM) = [ F(sample) - F(blank) x n ] / slope (uM)-1

n = sample dilution factor 

**note** if the calculated succinate concentration is >40uM for the Fluorometric assay, dilute the sample in ultrapure water and repeat the assay,
multipley by the dilution factor 

**important** in one of our succinate runs we diluted the sample that was already in the wells and recorded, The authors here seem to refer to the sample 
before all reagents are added and incubated to truly 'dilute' the signal, time will tell below whether this is the case because we ran it different ways 
in order to test! 

  * remember - we used ~10 mg tissue for each sample, keep this is mind when/of repeating this assay
  
```{r Correct for background}

## Objectve to subtract the blank from all standards and samples. 

# Plate 1 

Plate1.standards.sub <- Plate1.standards %>% 
                          dplyr::mutate(Abs_532nm_sub = 
                                          case_when(nmolMDA > 0 ~ Abs_532nm - Plate1.blanks.means$mean,
                                                    
                                                    nmolMDA == 0 ~ 0
                                                    
                                                    ))


Plate2.samples.sub <- Plate1.samples %>% 
                          dplyr::mutate(Abs_532nm_sub = Abs_532nm - Plate1.blanks.means$mean)


# plate 2


Plate2.standards.sub <-  Plate2.standards %>% 
                          dplyr::mutate(Abs_532nm_sub = 
                                          case_when(nmolMDA > 0 ~ Abs_532nm - Plate2.blanks.means$mean,
                                                    
                                                    nmolMDA == 0 ~ 0
                                                    
                                                    ))


Plate2.samples.sub <- Plate2.samples %>% 
                          dplyr::mutate(Abs_532nm_sub = Abs_532nm - Plate2.blanks.means$mean)


```


```{r Lipid peroxidation standard curves}
library('ggpmisc')

# Plate 1
Plate1_standard_curve <- Plate1.standards.sub %>% 
                                    dplyr::select(nmolMDA, Abs_532nm_sub) %>% 
                                    # dplyr::filter(Time_min %in% 30) %>% 
                                    dplyr::group_by(nmolMDA) %>% 
                                    dplyr::summarise(meanAbs = mean(Abs_532nm_sub),
                                                     sdAbs   = sd(Abs_532nm_sub),
                                                     n = n()) %>% 
                          
                          ggplot(aes(x = as.numeric(nmolMDA), y = as.numeric(meanAbs))) + 
                          geom_point() + 
                          stat_poly_line() +
                          stat_poly_eq(use_label(c("eq", "R2"))) +
                          theme_bw() + 
                          xlab('known uM MDA') +
                          ylab('Abs at 532 nm') 


# Plate 2
Plate2_standard_curve <- Plate2.standards.sub %>% 
                                    dplyr::select(nmolMDA, Abs_532nm_sub) %>% 
                                    # dplyr::filter(Time_min %in% 30) %>% 
                                    dplyr::group_by(nmolMDA) %>% 
                                    dplyr::summarise(meanAbs = mean(Abs_532nm_sub),
                                                     sdAbs   = sd(Abs_532nm_sub),
                                                     n = n()) %>% 
                          
                          ggplot(aes(x = as.numeric(nmolMDA), y = as.numeric(meanAbs))) + 
                          geom_point() + 
                          stat_poly_line() +
                          stat_poly_eq(use_label(c("eq", "R2"))) +
                          theme_bw() + 
                          xlab('known uM MDA') +
                          ylab('Abs at 532 nm')


# output 
pdf(paste("Output/Colorimetric_assays/Lipid_Peroxidation/Standard_Curve_Lipid_Peroxidation.pdf", sep =''), 
    width=10, 
    height=5)
ggarrange(Plate1_standard_curve,
  Plate2_standard_curve,
  nrow=2)
dev.off()

```

## Calculate lipid peroxidation 

* as written in the protocol...

1. Average the duplicate reading for each standard and sample.

*done in the chunk above for standards* 

*do this in the chunk below for the samples* 

2. Subtract the mean value of the blank (Standard #1) from all standards and sample
readings. This is the corrected absorbance.

*done in the chunks above in Plate1.samples.sub and Plate2.samples.sub* 

3. Plot the corrected values for each standard as a function of the final concentration of
MDA.

*done in the chunk above , plot to show the y = mx+ b equation* 

4. Draw the best smooth curve through these points to construct the standard curve. Most
plate reader software or Excel can plot these values and curve fit. Calculate the trendline
equation based on your standard curve data (use the equation that provides the most
accurate fit).

*done in the chunk above , plot to show the y = mx+ b equation* 

5. Apply the corrected sample OD reading to the standard curve to get MDA amount (y =
mx + b) in the sample wells.

*do this in the chunk below for the samples* 

6. Concentration of MDA in the test samples is calculated as:

  MDA Concentration = ( A / [mg or mL]) x 4 x D = nmol per ml or nmol mg
  
  A = Amount of MDA in sample calculated from the standard curve (nmol).*in other words solve for y from the stnadrd eq above*
  mg = Original tissue amount used (e.g. 10 mg). *we did not measure this, instead correct to total protein* 
  mL = Original plasma volume used (0.020 mL). *not applicable*
  4* = Correction for using 200 μL of the 800 μL Reaction Mix. *do this*
  D = Sample dilution factor if sample is diluted to fit within the standard curve range (prior
  to reaction well set up). *we did not dilute further in our runs, no need for D*

* **NOTE** we did not quantify the amount of starting tissue but instead we will correct to 
total protein of the sample tissue homogenate



```{r Calculate Lipid peroxidation}

library(tidyr)

# y = mx + b for plates 1 and 2 
# plate 1
m1 = 0.296
b1 = -0.0139
# plate 2
m2 = 0.297
b2 = -0.0169


# mean total protein
meanTP <- TotalProtein_final %>%  
                    group_by(Unique_ID) %>% 
                    dplyr::summarise(meanTotalProtein_ug_mL = mean(TotalProtein_ug_mL))
nrow(meanTP) # 29 sample for total protein, remeber we had a ton f samples above our absorbance threshold
length(unique(meanTP$Unique_ID)) # 29

# Raw Lipid per oxidation from y = mx + b alone

meanMDA_raw <- rbind(Plate1.samples, Plate2.samples) %>% 
                        dplyr::select(Plate, Temperature.Number, Tank.Number, Volume, Abs_532nm) %>% 
                        dplyr::mutate(Unique_ID = paste0('Tank',Tank.Number,'_', 'Temperature',  Temperature.Number)) %>% # unique ID t0 group by
                        dplyr::group_by(Plate, Temperature.Number, Tank.Number, Unique_ID) %>% 
                        dplyr::summarise(meanAbs = mean(Abs_532nm)) %>% 
                        dplyr::mutate(A = case_when(Plate == 1 ~ (m1*meanAbs) + b1, # y = mx + b
                                                    Plate == 2 ~ (m2*meanAbs) + b2),
                                      Treatment = case_when(Tank.Number %in% c(1:7) ~ 'Ambient',
                                                            Tank.Number %in% c(8:14) ~ 'Heated'))


MDA_per.ug.protein <- merge(meanMDA_raw, meanTP, 
                                  by = c('Unique_ID')) %>% 
                                # TotalProtein_ug is the amount of proetin in 25 ul of homogenate 
                                # correct for the volume in the Succiante sample, in all cases we used 20 ul of the homogenate
                                # to measure MDA
                                # (TotalProtein_ug_mL*(20/1000) - to 
                                dplyr::mutate(MDA_uM_ug_protein = (A/ (meanTotalProtein_ug_mL*(200/1000)) ) * 4) %>% 
                                # filter(!Cs_Activty_per_ug_protein < 0) %>% 
                                dplyr::select(Temperature.Number,
                                              Tank.Number,
                                              Treatment,
                                              Unique_ID,
                                              meanAbs,
                                              A,
                                              meanTotalProtein_ug_mL,
                                              MDA_uM_ug_protein
                                              ) #%>% 
                                # dplyr::group_by(Unique_ID, Temperature.Number, Tank.Number, Treatment) %>% 
                                # dplyr::summarise(meanTotalProtein_ug_mL = mean(TotalProtein_ug_mL),n= n(),
                                                 # MDA_uM_ug_protein = median(MDA_uM_ug_protein)) 
nrow(MDA_per.ug.protein) # 29


### master table 
Master_Table <- MDA_per.ug.protein %>% 
                  # dplyr::filter(!meanCs_Activity_ug_protein < 0) %>% 
                  dplyr::mutate(Temperature = case_when(
                    Temperature.Number == 1 & Treatment %in% 'Heated' ~ "22",
                    Temperature.Number == 2 & Treatment %in% 'Heated' ~ "26",
                    Temperature.Number == 3 & Treatment %in% 'Heated' ~ "30",
                    Temperature.Number == 4 & Treatment %in% 'Heated' ~ "34",
                    Temperature.Number == 5 & Treatment %in% 'Heated' ~ "36",
                 
                    Temperature.Number == 1 & Treatment %in% 'Ambient' ~ "22_t1",
                    Temperature.Number == 2 & Treatment %in% 'Ambient' ~ "22_t2",
                    Temperature.Number == 3 & Treatment %in% 'Ambient' ~ "22_t3",
                    Temperature.Number == 4 & Treatment %in% 'Ambient' ~ "22_t4",
                    Temperature.Number == 5 & Treatment %in% 'Ambient' ~ "22_t5",
                  )) %>% 
                  dplyr::rename(Timepoint = Temperature.Number)

# write csv
write.csv(Master_Table, file = "Output/Colorimetric_assays/Lipid_Peroxidation/Calc_Master_Lipid_Peroxidation.csv")



```


```{r plot data}
# plot all data above 0
Plot_MeanSE <- Master_Table %>% 
                  # dplyr::filter(!MDA_uM_ug_protein > 350) %>%
                  # dplyr::filter(!meanSuccinate_per_ug_protein_10xpre > 600) %>%
                  ggplot(aes(x = as.factor(Temperature), 
                             y = MDA_uM_ug_protein,
                             # y = meanSuccinate_per_ug_protein_10xpre,
                             color=Treatment,
                             group=Treatment)) +
                         scale_colour_manual(breaks=c("Ambient", "Heated"), 
                                                             values=c("forestgreen","orange")) +
                         geom_point(aes(colour = Treatment), 
                                        position = position_dodge2(width = 0.4)) + 
                         stat_summary(fun.y="mean", size = 0.8, color = "black",
                                      position = position_dodge2(width = 0.4)) +
                         stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)), 
                                      fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                      geom = 'errorbar', width = 0.25, size = 1,
                                      position = position_dodge2(width = 10.4)) +
                         labs(title="uM MDA ug total protein", 
                              x ="Temperature", 
                              y = expression("MDA"~(~uM^{-1}*~ug~protein^{-1}))) + #"CS activity per g protein") +
                        # scale_x_discrete(labels=c("22", "26", "30" , "34", "36")) +
                         # scale_y_continuous(expand = c(0, 0), limits = c(0, 0.00035), 
                                            # breaks = seq(0, 0.00035, by = 0.00007)) +
                         theme_classic() +
                         theme(panel.grid.major = element_blank(),  
                               panel.grid.minor = element_blank(), 
                               axis.title =element_text(size=12),
                               axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                               axis.text=element_text(size=10),
                               plot.title = element_text(size=12),
                               legend.position="none") + 
                        facet_wrap(~Treatment, scales= "free_x")

# plot all data above 0 - omit outliers and present only the heated treatment
Plot_MeanSE_ambient <- Master_Table %>% 
                  dplyr::filter(!(Treatment %in% 'Heated'))%>%
                  ggplot(aes(x = as.factor(Temperature),
                             y = MDA_uM_ug_protein,
                             color=Treatment)) +
                         scale_colour_manual(breaks=("Ambient"),
                                             values=("forestgreen")) +
                         geom_point(aes(colour = Treatment),
                                        position = position_dodge2(width = 0.4)) +
                         stat_summary(fun.y="mean", size = 0.8, color = "black",
                                      position = position_dodge2(width = 0.4)) +
                         stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)),
                                      fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                      geom = 'errorbar', width = 0.25, size = 1,
                                      position = position_dodge2(width = 10.4)) +
                         labs(title="uM MDA ug total protein", 
                              x ="Temperature",
                              y = expression("MDA"~(~uM^{-1}*~ug~protein^{-1}))) + #"CS activity per g protein") +
                        # scale_x_discrete(labels=c("22", "26", "30" , "34", "36")) +
                         # scale_y_continuous(expand = c(0, 0), limits = c(0, 0.00035),
                         #                    breaks = seq(0, 0.00035, by = 0.00007)) +
                         theme_classic() +
                         theme(panel.grid.major = element_blank(),
                               panel.grid.minor = element_blank(),
                               axis.title =element_text(size=12),
                               axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                               axis.text=element_text(size=10),
                               plot.title = element_text(size=12),
                               legend.position="none")

# Plot_MeanSE_ambient

# plot all data above 0 - omit outliers and present only the heated treatment
Plot_MeanSE_heated <- Master_Table %>% 
                  dplyr::filter(!(Treatment %in% 'Ambient')) %>%
                  ggplot(aes(x = as.factor(Temperature),
                             y = MDA_uM_ug_protein,
                             color=Treatment)) +
                         scale_colour_manual(breaks=("Heated"),
                                             values=("orange")) +
                         geom_point(aes(colour = Treatment),
                                        position = position_dodge2(width = 0.4)) +
                         stat_summary(fun.y="mean", size = 0.8, color = "black",
                                      position = position_dodge2(width = 0.4)) +
                         stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)),
                                      fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                      geom = 'errorbar', width = 0.25, size = 1,
                                      position = position_dodge2(width = 10.4)) +
                         labs(title="uM MDA ug total protein", 
                              x ="Temperature",
                              y = expression("MDA"~(~uM^{-1}*~ug~protein^{-1}))) + #"CS activity per g protein") +
                        # scale_x_discrete(labels=c("22", "26", "30" , "34", "36")) +
                         # scale_y_continuous(expand = c(0, 0), limits = c(0, 0.00035),
                         #                    breaks = seq(0, 0.00035, by = 0.00007)) +
                         theme_classic() +
                         theme(panel.grid.major = element_blank(),
                               panel.grid.minor = element_blank(),
                               axis.title =element_text(size=12),
                               axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                               axis.text=element_text(size=10),
                               plot.title = element_text(size=12),
                               legend.position="none")


ggarrange(Plot_MeanSE_ambient, Plot_MeanSE_heated)

# output 
pdf(paste("Output/Colorimetric_assays/Lipid_Peroxidation/Lipid_Peroxidation_AmbinetvHeated.pdf", sep =''), 
    width=10, 
    height=4)
ggarrange(Plot_MeanSE_ambient, Plot_MeanSE_heated)
dev.off()


# pdf(paste("Output/Colorimetric_assays/Citrate_Synthase/Citrate_Synthase_Heated.pdf", sep =''), 
#     width=5, 
#     height=4)
# print(Plot_MeanSE_heated)
# dev.off()
# 
# pdf(paste("Output/Colorimetric_assays/Citrate_Synthase/Citrate_Synthase_Ambient.pdf", sep =''), 
#     width=5, 
#     height=4)
# print(Plot_MeanSE_ambient)
# dev.off()

```


# STATISTICS


```{r run stats}
library(rstatix)
Master <- rbind(Plate_1_CS_per.g.protein, Plate_2_CS_per.g.protein) %>% 
                  dplyr::filter(!(meanCs_Activty < 0  | meanCs_Activty > 0.0006)) %>%
                  dplyr::mutate(Temperature = case_when(
                    Temperature.Number == 1 ~ "22",
                    Temperature.Number == 2 ~ "26",
                    Temperature.Number == 3 ~ "30",
                    Temperature.Number == 4 ~ "34",
                    Temperature.Number == 5 ~ "38",
                  ))


Master %>% 
            group_by(Temperature) %>%
            anova_test(dv = meanCs_Activty, between = Treatment) %>%
            get_anova_table() %>%
            adjust_pvalue(method = "bonferroni")
  
```

