---
title: "PEPCK.Rmd"
author: "Sam Gurr"
date: "2025-1-7"
output: html_document
---

-   Last updates: 1 7, 2025

# Thermal Performacne (pilot test) & INFISH projet with Indya Limpkin & Ashley Hallengren

## Ran a PEPCK on C virginica gill tissue homogenate, corrected to

total protein using pierce BCA protein assay (albumin standard)

### SET UP

* note: add your own directory below to navigate to RAnalysis folder containing the data, output, and script subfolders 
that reflect the google drive. alterntively clone the repository EAD-ASEB-Cvirginica_Thermal_Performance from github
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
# knitr::opts_knit$set(root.dir = "C:/Users/katherine.mcfarland/Documents/GitHub/EAD-ASEB-Cvirginica_Thermal_Performance/larvae") # Katie's
knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Cvirginica_Thermal_Performance/RAnalysis") # Sam's
#knitr::opts_knit$set(root.dir = "C:/Users/samuel.gurr/Documents/Github_repositories/EAD-ASEB-Cvirginica_Thermal_Performance/RAnalysis") # Sam's work

```


## Load packages 

* use install.packges('pacakge name') if you do not have these downloaded
```{r setup, include=TRUE}

library(ggpubr)
library(ggplot2)
library(dplyr)
library(lmtest) # to receive p value from betareg model
# library(FSA) # for the Dun test post hoc for SRH non-parametric 2 way anova]
library(emmeans)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(car)
library(lmerTest)
library(tidyr)
library(reshape2)
library(ggpubr)
library(nlme)
library(rcompanion) # to run the Schrier -Ray-Hare non parametric 2 way 
library(ggpmisc) # stat_poly for inserting equation and R2 for ggplot line 
```

## Function to load files

-   At your service, from 'spec to table' ...I'm so funny

```{r Spec_to_table}

Spec_to_table <- function(datapath, platenumber, skipnrows, daterecorded) {
  
# import the data in raw text form exported from the spectrophotometer
        rawtextfile <- read.delim(datapath,
                              header = F,
                              skip=skipnrows,
                              fileEncoding="latin1")
  

        rawmatrix   <- as.matrix(rawtextfile[c(1:8),c(3:14)])
        colnames(rawmatrix) = c("1","2","3","4","5","6","7","8","9","10","11","12")
        rownames(rawmatrix) = c("A","B","C","D","E","F","G","H")
        
        raw_table <- as.data.frame.table(rawmatrix, responseName = "value") %>% 
                                    dplyr::rename(well_row=Var1, well_column=Var2, Abs_raw=value) %>% 
                                    dplyr::mutate(well=paste0(well_row,well_column),
                                                  Run_date =daterecorded,
                                                  Plate=platenumber) %>% 
                                    dplyr::select(-c(well_row,well_column))
        # reorder the columns because I prefer it this way..
        
        raw_table_ordered <- raw_table[,c(3,4,2,1)]
}


```

## LOAD DATA

-   Use the spec_to_table to import raw spec txt files and compile to a table, yes that is what it does

-   add an "about" column - we have a few hiccups with this assay and reran it with dilutions

  - no dilution = ran without additional dilutions 
  - 5x dilution post = after we ran the no dilution we saw any values were ut of range (too high) we diluted 5x with ultrapure water direct to the plate and reran
  - 5x dilution post 30 mins = same as the 5 x dilution but after a 30 minute incubation we ran it again

```{r load Succinate Synthase raw data and build master files}

# plate 1
PEPCK_Plate1_0mins  <- Spec_to_table('Data/PEPCK/20241114_PEPCK_Kinetic_Plate1_20mins.txt', 1, 3, 20241114) %>% 
                            dplyr::mutate(Time = 0)
PEPCK_Plate1_5mins  <- Spec_to_table('Data/PEPCK/20241114_PEPCK_Kinetic_Plate1_20mins.txt', 1, 12, 20241114) %>% 
                            dplyr::mutate(Time = 5)
PEPCK_Plate1_10mins <- Spec_to_table('Data/PEPCK/20241114_PEPCK_Kinetic_Plate1_20mins.txt', 1, 21, 20241114) %>% 
                            dplyr::mutate(Time = 10)
PEPCK_Plate1_15mins <- Spec_to_table('Data/PEPCK/20241114_PEPCK_Kinetic_Plate1_20mins.txt', 1, 30, 20241114) %>% 
                            dplyr::mutate(Time = 15)
PEPCK_Plate1_20mins <- Spec_to_table('Data/PEPCK/20241114_PEPCK_Kinetic_Plate1_20mins.txt', 1, 39, 20241114) %>% 
                            dplyr::mutate(Time = 20)

PEPCK_Plate1_20mins_stand <- Spec_to_table('Data/PEPCK/20241114_PEPCK_Endpoint_Plate1_20mins_standards.txt', 1, 3, 20241114) %>% 
                                  dplyr::mutate(Time = 'standard')

PEPCK_Plate1_MASTER <- rbind(
                             PEPCK_Plate1_0mins,
                             PEPCK_Plate1_5mins,
                             PEPCK_Plate1_10mins,
                             PEPCK_Plate1_15mins,
                             PEPCK_Plate1_20mins,
                             PEPCK_Plate1_20mins_stand
                             )

# plate 2
PEPCK_Plate2_0mins  <- Spec_to_table('Data/PEPCK/20241114_PEPCK_Kinetic_Plate2_20mins.txt', 2, 3, 20241114) %>% 
                            dplyr::mutate(Time = 0)
PEPCK_Plate2_5mins  <- Spec_to_table('Data/PEPCK/20241114_PEPCK_Kinetic_Plate2_20mins.txt', 2, 12, 20241114) %>% 
                            dplyr::mutate(Time = 5)
PEPCK_Plate2_10mins <- Spec_to_table('Data/PEPCK/20241114_PEPCK_Kinetic_Plate2_20mins.txt', 2, 21, 20241114) %>% 
                            dplyr::mutate(Time = 10)
PEPCK_Plate2_15mins <- Spec_to_table('Data/PEPCK/20241114_PEPCK_Kinetic_Plate2_20mins.txt', 2, 30, 20241114) %>% 
                            dplyr::mutate(Time = 15)
PEPCK_Plate2_20mins <- Spec_to_table('Data/PEPCK/20241114_PEPCK_Kinetic_Plate2_20mins.txt', 2, 39, 20241114) %>% 
                            dplyr::mutate(Time = 20)

PEPCK_Plate2_20mins_stand <- Spec_to_table('Data/PEPCK/20241114_PEPCK_Endpoint_Plate2_20mins_standards.txt', 2, 3, 20241114)%>% 
                                  dplyr::mutate(Time = 'standard')

PEPCK_Plate2_MASTER <- rbind(
                             PEPCK_Plate2_0mins,
                             PEPCK_Plate2_5mins,
                             PEPCK_Plate2_10mins,
                             PEPCK_Plate2_15mins,
                             PEPCK_Plate2_20mins,
                             PEPCK_Plate2_20mins_stand
                             )

```

```{r load Total protein correction data for succinate}

# plate 1 
TP_Plate1_master <- Spec_to_table('Data/Total_Protein_PEPCK/20241115_TotalProteinPEPCK_Plate1_590nm.txt', 1, 3, 20241115) %>% 
                        dplyr::rename(Abs_590nm = Abs_raw)

# plate 2
TP_Plate2_master <- Spec_to_table('Data/Total_Protein_PEPCK/20241115_TotalProteinPEPCK_Plate2_590nm.txt', 2, 3, 20241115) %>% 
                        dplyr::rename(Abs_590nm = Abs_raw) 

```


```{r load reference metadata and call target datasets}

metadata <- read.csv('Data/PEPCK_metadata.csv', sep = ",", header=T)

# rename a few columns so they can merge with the data (in next chunk)

# Plate 1 
metadata.PEPCK_Plate_1 <- metadata %>% 
                          dplyr::filter(Assay.Type %in% "PEPCK" & Plate.Number == 1) %>% 
                          dplyr::rename(Plate = Plate.Number,
                                        well  = Well.ID)

# Plate 2
metadata.PEPCK_Plate_2 <- metadata %>% 
                        dplyr::filter(Assay.Type %in% "PEPCK" & Plate.Number == 2) %>% 
                        dplyr::rename(Plate = Plate.Number,
                                      well  = Well.ID)



# Plate 1 
metadata.TP_Plate_1 <- metadata %>% 
                        dplyr::filter(Assay.Type %in% "Total Protein" & Plate.Number == 1) %>% 
                        dplyr::rename(Plate = Plate.Number,
                                      well  = Well.ID)
# Plate 2
metadata.TP_Plate_2 <- metadata %>% 
                        dplyr::filter(Assay.Type %in% "Total Protein" & Plate.Number == 2) %>% 
                        dplyr::rename(Plate = Plate.Number,
                                      well  = Well.ID)

```

## PREPARE DATA:

* merge metadata with data 
```{r merge metadata with data}
# View(metadata.PEPCK_Plate_2)
# MDA
# Plate 1 
Plate1 <- merge(metadata.PEPCK_Plate_1, PEPCK_Plate1_MASTER, by = c('Plate', 'well'))                          
nrow(PEPCK_Plate1_MASTER) == nrow(Plate1) # must be TRUE

# Plate 2
Plate2 <- merge(metadata.PEPCK_Plate_2, PEPCK_Plate2_MASTER, by = c('Plate', 'well'))                          
nrow(PEPCK_Plate2_MASTER) == nrow(Plate2) # must be TRUE - NOT HERE THOUGH because metadtata did not include the empty IDs, we have 78 with sample, all good


# Total protein
# Plate 1 
Plate1.TP <- merge(metadata.TP_Plate_1, TP_Plate1_master, by = c('Plate', 'well'))                          
nrow(TP_Plate1_master) == nrow(Plate1.TP) # must be TRUE
Plate1.TP <- subset(Plate1.TP, !is.na(Sample.Type)) # 96 rows

# Plate 2 
Plate2.TP <- merge(metadata.TP_Plate_2, TP_Plate2_master, by = c('Plate', 'well'))                          
nrow(TP_Plate2_master) == nrow(Plate2.TP) # must be TRUE - NOT HERE THOUGH 
Plate2.TP <- subset(Plate2.TP, !is.na(Sample.Type)) # now 80 total rows with NAs omitted (not sample, blank nor stnadard)


```




* subset standards, blanks, and samples
```{r subset data types}

# PEPCK
# Plate 1 
Plate1.samples   <- Plate1 %>% dplyr::filter(Sample.Type %in% c('Sample', 'SBC')) %>% dplyr::rename(Abs_570nm = Abs_raw)
# Plate1.blanks    <- Plate1 %>% dplyr::filter(Sample.Type %in% 'Blank') # space, why... whatevs
Plate1.standards <- Plate1 %>% dplyr::filter(Sample.Type %in% c('Blank','Standard')) %>% 
                               dplyr::rename(Abs_570nm = Abs_raw,
                                             nmolPyruvate = Standards)

# Plate 2
Plate2.samples   <- Plate2 %>% dplyr::filter(Sample.Type %in% c('Sample', 'SBC')) %>% dplyr::rename(Abs_570nm = Abs_raw)
# Plate2.blanks    <- Plate2 %>% dplyr::filter(Sample.Type %in% 'Blank')  # space, why... whatevs
Plate2.standards <- Plate2 %>% dplyr::filter(Sample.Type %in% c('Blank','Standard')) %>% 
                               dplyr::rename(Abs_570nm = Abs_raw,
                                             nmolPyruvate = Standards)

# Total protein
# Plate 1 
Plate1_TP.samples    <- Plate1.TP %>% dplyr::filter(Sample.Type %in% 'Sample')
Plate1_TP.blanks     <- Plate1.TP %>% dplyr::filter(Sample.Type %in% 'Blank')
Plate1_TP.standards  <- Plate1.TP %>% dplyr::filter(Sample.Type %in% c('Standard', 'Blank'))

(nrow(Plate1_TP.samples) + 
    # nrow(Plate1_TP.blanks) + 
      nrow(Plate1_TP.standards)) == nrow(Plate1.TP) # sanity check, must be true


# Plate 2 
# NOTE What hte hell is 'Standard_AssayBufferTest'??? 
# we ran multiple samples as extras with the assay buffer as diluent to make sure whether 
# the same solution used in the standards does not confound the signal for these samples!
Plate2_TP.samples    <- Plate2.TP %>% dplyr::filter(Sample.Type %in% c('Sample', 'Standard_AssayBufferTest'))
Plate2_TP.blanks     <- Plate2.TP %>% dplyr::filter(Sample.Type %in% 'Blank') 
Plate2_TP.standards  <- Plate2.TP %>% dplyr::filter(Sample.Type %in% c('Standard', 'Blank'))


(nrow(Plate2_TP.samples) + 
    # nrow(Plate2_TP.blanks) + 
      nrow(Plate2_TP.standards)) == nrow(Plate2.TP) # sanity check, must be true

```


## CALCULATIONS FOR TOTAL PROTEIN

```{r total protein standard curve}

# assign known BCA values to standard IDs A-I
BCA_standards <- rbind(Plate1_TP.standards, Plate2_TP.standards) %>%  
  dplyr::rename(Abs_562nm = Abs_590nm) %>% 
  dplyr::mutate(BCA_ug_mL = case_when(Standards %in% 'A' ~ 2000,
                                      Standards %in% 'B' ~ 1500,
                                      Standards %in% 'C' ~ 1000,
                                      Standards %in% 'D' ~ 750,
                                      Standards %in% 'E' ~ 500,
                                      Standards %in% 'F' ~ 250,
                                      Standards %in% 'G' ~ 125,
                                      Standards %in% 'H' ~ 25,
                                      Standards %in% 'I' ~ 0)) %>% 
  dplyr::mutate(Abs_562nm = as.numeric(Abs_562nm)) %>% 
  dplyr::select(Plate, Standards, BCA_ug_mL, Abs_562nm)

# Run standard curve, calculate totalprotein 
BCA_background_zero <- BCA_standards %>% 
                        dplyr::filter(Standards %in% 'I') %>% # the zero standard
                        # dplyr::group_by(Plate, Standards, BCA_ug_mL) %>% # group by to get the means
                        Rmisc::summarySE(groupvars  = c('Plate', 'Standards', 'BCA_ug_mL'),
                                         measurevar = 'Abs_562nm') # get all the stats 


# Plate 1, blank to correct by is 0.09895
# Plate 2, blank to correct by is 0.09710

# Absorbance corrected - take the mean of any duplicates
BCA_standards_means <- BCA_standards %>% 
                        dplyr::filter(!BCA_ug_mL %in% 0) %>% 
                        dplyr::mutate(Abs_562nm_cor = 
                                      as.numeric(
                                        case_when(Plate == 1 ~ (Abs_562nm-0.09895),
                                                Plate == 2 ~ (Abs_562nm-0.09710) ) ) ) %>% 
                        dplyr::select(-Abs_562nm) %>% 
                        Rmisc::summarySE(groupvars  = c('Plate', 'Standards', 'BCA_ug_mL'),
                                         measurevar = 'Abs_562nm_cor') %>% # get all the stats 
                        dplyr::rename(mean = Abs_562nm_cor)

# plot it insert the quadratic formaula using ggpmisc
BCA_stand_plots_quadratic <- BCA_standards_means %>% 
                    # QUADRATIC SMOOTH LINE WORKS BEST HERE (MANUFACTURERS INSTRUCTIONS)
                     # dplyr::filter(!(Plate %in% 2 & Standards %in% 'D')) %>% # hash me out to test
                     ggplot(aes(y=mean, x=BCA_ug_mL)) + 
                        geom_point() +
                        theme_bw() +
                        labs(y= "Net Abs 562nm", x = "Protein Concentration in ug/mL") +
                        #geom_line() +
                        #stat_poly_line(color='red') +
                        #geom_smooth() +
                        stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1) +
                        stat_poly_eq(parse=T, aes(label = ..eq.label..), formula=y ~ x + I(x^2)) +
                        ggtitle('Total protein: Quadratic curve') +
                        #stat_poly_eq(use_label(c("eq", "R2"))) +
                        facet_wrap(~Plate) 

# output 
pdf(paste("Output/Colorimetric_assays/Lipid_Peroxidation/Standard_Curve_BCA_MDAcorrection.pdf", sep =''), 
    width=10, 
    height=7)
print(ggarrange(BCA_stand_plots_quadratic))
dev.off()

```


* Use BCA standard curves to calculate TP 

```{r calc TP}
library(grDevices)
# Standard curve, Plate 1 equation y = 0.0195 + 0.000972x - 1.05x10^-7x^2 - need to solve for x!
# Standard curve, Plate 2 equation y = 0.0196 + 0.000751x - 3.22x10^-8x^2 - need to solve for x!

# Standard curve, Plate 1
a1 <- -1.05*10^-7
b1 <- 0.000972
c1 <- 0.0195
# EQ: (-(b1) + sqrt( (b1^2) - (4*(((a1)-Abs_562nm_cor))*(c1)) ))/(2*a1)

# Standard curve, Plate 2
a2 <- -3.22^-8
b2 <- 0.000751
c2 <- 0.0196
# EQ: (-(b2) + sqrt( (b2^2) - (4*a2*(c2-Abs_562nm_cor)) ) ) / (2*a2)


# linear equation plate 1 == (Abs_562nm_cor - 0.0195)/0.000972
# linear equation plate 2 == (Abs_562nm_cor - 0.0196)/0.000751


# IMPORTANT! we used 25 ul of the standards and 25 ul of the unknowns (samples) 
# therefore we can interpret the unknown direct to the the standard curve without having 
# to account for addition factors, fot example, if we used 5 ul unknown (sample) we would have to adjust 
# by multiplying by 5 to reach the standard curve 

V = 0.025 # 25 ul or 0.025 mL

# Again, remember Plate 1, blank to correct by is 0.09895
# Again, remember Plate 2, blank to correct by is 0.09710

# Sanity check Lets look at the absorbance vs. totla protein concentration data 

TotalProtein_final <- rbind(Plate1_TP.samples, Plate2_TP.samples) %>% 
                      dplyr::rename(Abs_562nm = Abs_590nm) %>% 
                      dplyr::mutate(Abs_562nm = as.numeric(Abs_562nm)) %>% 
                      # dplyr::filter(!Abs_562nm >= 4) %>% 
                      dplyr::select(Plate,Tank.Number,Temperature.Number,Abs_562nm) %>% 
                      # dplyr::filter(!Abs_562nm > 3.99) %>% # data as 4.00 is above the detection limit, omit
                      dplyr::mutate(Unique_ID = 
                                      paste0('Plate_',Plate,'Tank',Tank.Number, '_',
                                             'Temperature', Temperature.Number)) %>% # unique ID t0 group by
                      dplyr::mutate(Abs_562nm_cor = # correct the raw abs, subtract background
                                   case_when(Plate == 1 ~ (Abs_562nm-0.09895), # for plate 1
                                             Plate == 2 ~ (Abs_562nm-0.09710) ) ) %>% # for plate 2 
                      dplyr::mutate(TotalProtein_ug_mL = 
                                    case_when(
                                      # linear fr neg discrim. - luckily only two values from plate 2
                                      # Scallop_ID %in% c(33, 51) ~ 
                                      #   ((Abs_562nm_cor - 0.224)/0.000911),
                                      # quadratic for Plate 1
                                      Plate == 1 ~ 
                                        ((-(b1) + sqrt( (b1^2) - (4*a1*(c1-Abs_562nm_cor)) ) ) / (2*a1)), 
                                      # quadratic for plate 2
                                      Plate == 2 ~ 
                                        # ((-(b2) + sqrt( (b2^2) - (4*a2*(c2-Abs_562nm_cor)) ) ) / (2*a2)) ),
                                        (Abs_562nm_cor - 0.0196)/0.000751),
                                    # ug per mL concentration to ug in 25 ul sample 
                                    TotalProtein_ug = TotalProtein_ug_mL*V) %>% 
                      na.omit()

length(unique(TotalProtein_final$Unique_ID)) # 39 samples retained for total protein

# View(TotalProtein_final)
nrow(TotalProtein_final) # 78

# NOTE! I found that some of my discriminants are negative (b^2 - 4ac) IN PLATE 2!! (as 'NaN)
# so Im going to extrapolate from a linear regression curve based on the final samples 

# first lets plot absorbance and actual concentration calacuated 
calc_BCA_plot <- TotalProtein_final %>% 
                    ggplot(aes(y = Abs_562nm_cor, 
                               x  = TotalProtein_ug_mL)) +
                          geom_point() + 
                    theme_bw() + 
                    ggtitle('Total protein: Calculated BCA by Net Absorbance') +
                    facet_wrap(~Plate)
calc_BCA_plot 

#print
pdf(paste("Output/Colorimetric_assays/PEPCK/Standard_Curve_BCA_MDAcorrection.pdf", sep =''), 
    width=10, 
    height=14)
ggpubr::ggarrange(BCA_stand_plots_quadratic,
          calc_BCA_plot,
          nrow = 2)
dev.off()
# write csv
write.csv(TotalProtein_final, file = "Output/Colorimetric_assays/PEPCK/TotalProteinBCA_PEPCKcorrection.csv")


```

## CALCULATIONS FOR PEPCK:


1. Subtract the 0 Standard A570 reading 
from all A570 readings, including 
Sample(s) (S), Sample Background 
Control (SBC), and Positive Controls.

**NOTE** the pyruvate standard curve was measured as end point mode after the last reading and end of incubation time
(all other reading in kinetic) - the end point standard mode data is called 'standard' in the 'Time' column


```{r}

# Run standard curve, calculate totalprotein 
Plate1_StandardA570_zero <- Plate1.standards %>% 
                              dplyr::filter(Sample.Type %in% 'Blank' & 
                                              Time %in% 'standard') %>% # the zero standard
                              # dplyr::group_by(Plate, Standards, BCA_ug_mL) %>% # group by to get the means
                              Rmisc::summarySE(groupvars  = c('Plate', 'nmolPyruvate'),
                                               measurevar = 'Abs_570nm') # get all the stats 



Plate2_StandardA570_zero <- Plate2.standards %>% 
                              dplyr::filter(Sample.Type %in% 'Blank' & 
                                              Time %in% 'standard') %>% # the zero standard
                              # dplyr::group_by(Plate, Standards, BCA_ug_mL) %>% # group by to get the means
                              Rmisc::summarySE(groupvars  = c('Plate', 'nmolPyruvate'),
                                               measurevar = 'Abs_570nm') # get all the stats 



Plate1.samples$Abs_570nm_blankcor <- Plate1.samples$Abs_570nm - Plate1_StandardA570_zero$Abs_570nm
Plate2.samples$Abs_570nm_blankcor <- Plate2.samples$Abs_570nm - Plate2_StandardA570_zero$Abs_570nm
```

2. Plot the Pyruvate standard curve.  

```{r}

PEPCK_stand_curve <- rbind(Plate1.standards, Plate2.standards) %>% 
                                dplyr::filter(Time %in% 'standard') %>% 
                                dplyr::mutate(Abs_570nm_blankcor = 
                                              case_when(
                                                Sample.Type %in% 'Blank' ~ 0,
                                                (Sample.Type %in% 'Standard' & Plate == 1) ~ 
                                                        Abs_570nm - Plate1_StandardA570_zero$Abs_570nm,
                                                (Sample.Type %in% 'Standard' & Plate == 2)  ~ 
                                                        Abs_570nm - Plate2_StandardA570_zero$Abs_570nm
                                              )) %>%
                      Rmisc::summarySE(measurevar = 'Abs_570nm_blankcor', 
                                       groupvars = c('Plate','nmolPyruvate')) %>% 
                      dplyr::filter(!(Plate == 2 & nmolPyruvate == 2)) %>% 
                    # QUADRATIC SMOOTH LINE WORKS BEST HERE (MANUFACTURERS INSTRUCTIONS)
                     ggplot(aes(y=Abs_570nm_blankcor, x=as.numeric(nmolPyruvate))) + 
                        geom_point() +
                        theme_bw() +
                        labs(y= "Abs 570nm", x = "nmol Pyruvate standard") +
                        #geom_line() +
                        #stat_poly_line(color='red') +
                        #geom_smooth() +
                        stat_smooth(method = "lm", formula = y ~ x, size = 1) +
                        stat_poly_eq(parse=T, aes(label = ..eq.label..), formula=y ~ x) +
                        ggtitle('PEPCK Standard: Linear curve') +
                        #stat_poly_eq(use_label(c("eq", "R2"))) +
                        facet_wrap(~Plate) 

PEPCK_stand_curve
```

3. Choose two time points (T1 and T2) in the 
linear range to calculate PEPCK Activity 
in Sample(s) (S). 

4. If the Sample Background Control (SBC) 
A570 readings are high, subtract the 
Sample Background Control (SBC) 
A570 reading from its paired Sample 
(S) A570 reading.  

5. Calculate ΔACorrected for each Sample. 
ΔACorrected = (AbsT2_S – AbsT1_S) - (AbsbsT2_SBC - AbsT1_SBC)  

```{r}

PEPCK_ΔACorrected <- rbind(Plate1.samples,Plate2.samples) %>% 
                                dplyr::select(Plate,Time,Sample.Type,
                                              Volume,Temperature.Number,
                                              Tank.Number,Abs_570nm_blankcor) %>% 
                                # 3
                                dplyr::filter(Time %in% c(10,20))  %>% 
                                pivot_wider(names_from = Time,
                                            values_from = Abs_570nm_blankcor) %>% 
                                # 4, not necessary
                                # 5
                                dplyr::mutate(ΔACorrected = `20` - `10`)
  
  
```

6. Apply ΔACorrected to the Pyruvate Standard 
curve to get B nmol of Pyruvate 
generated by PEPCK at the reaction time 
(ΔT = T2 - T1). 

7. Calculate the PEPCK activity of the 
Sample: 

PEPCK Activity (nmol/min/L or mU/L) =  [B/(ΔT × V)] × D 

  - B =  Pyruvate amount from Pyruvate Standard curve (pmol) 
  - ΔT =  Reaction time (T2 - T1) (minutes) 
  - V =  Sample volume added into the reaction well (L) 
  - D =  Sample Dilution factor (if applicable; 
  - D = 1 for undiluted Samples)

```{r}


PEPCK_Calc <- merge(PEPCK_ΔACorrected, 
                    (TotalProtein_final %>% dplyr::select(Plate,Temperature.Number,Tank.Number,TotalProtein_ug)), 
                    by = c('Plate','Temperature.Number','Tank.Number')) %>% 
                        # 6
                        dplyr::mutate(ΔACorrected_stand = 
                                      case_when(Plate == 1 ~ (ΔACorrected-0.0565)/(0.123),
                                                Plate == 2 ~ (ΔACorrected-0.000603)/(0.133) )) %>% 
                        dplyr::filter(!Sample.Type %in% 'SBC') %>% 
                        # 7
                        dplyr::mutate(
                          PEPCK_Activity_nmolPyruvante_ngProtein = 
                          (
                            ((ΔACorrected_stand / (10 * Volume)) * 1 )
                               / (TotalProtein_ug/1000)
                                    )
                        )



Master_Table <- PEPCK_Calc %>% 
                  filter(Volume == 30) %>% # hash this out to take a mean of all of them, we used 10 and 30 ul volume
                  dplyr::mutate(Treatment = case_when(
                                               Tank.Number %in% c(1:7) ~ 'Ambient',
                                               Tank.Number %in% c(8:14) ~ 'Heated'),
                               Temperature = case_when(
                                               Temperature.Number == 1 & Treatment %in% 'Heated' ~ "22",
                                               Temperature.Number == 2 & Treatment %in% 'Heated' ~ "26",
                                               Temperature.Number == 3 & Treatment %in% 'Heated' ~ "30",
                                               Temperature.Number == 4 & Treatment %in% 'Heated' ~ "34",
                                               Temperature.Number == 5 & Treatment %in% 'Heated' ~ "36",
                                             
                                               Temperature.Number == 1 & Treatment %in% 'Ambient' ~ "22_t1",
                                               Temperature.Number == 2 & Treatment %in% 'Ambient' ~ "22_t2",
                                               Temperature.Number == 3 & Treatment %in% 'Ambient' ~ "22_t3",
                                               Temperature.Number == 4 & Treatment %in% 'Ambient' ~ "22_t4",
                                               Temperature.Number == 5 & Treatment %in% 'Ambient' ~ "22_t5",
                                              )) %>% 
                  dplyr::rename(Timepoint = Temperature.Number) %>% 
                  # dplyr::group_by(Plate, Timepoint, Tank.Number, Treatment, Temperature) %>% 
                  Rmisc::summarySE(measurevar = 'PEPCK_Activity_nmolPyruvante_ngProtein', 
                                   groupvars = c('Plate', 'Timepoint', 'Tank.Number', 'Treatment', 'Temperature'))

# write csv
write.csv(Master_Table, file = "Output/Colorimetric_assays/PEPCK/Calc_Master_PEPCK.csv")

                  
                  
```

 






```{r plot data}
# plot all data above 0
Plot_MeanSE <- Master_Table %>% 
                  dplyr::filter(!PEPCK_Activity_nmolPyruvante_ngProtein > 1.7) %>%
                  # dplyr::filter(!meanSuccinate_per_ug_protein_10xpre > 600) %>%
                  ggplot(aes(x = as.factor(Temperature), 
                             y = PEPCK_Activity_nmolPyruvante_ngProtein,
                             # y = meanSuccinate_per_ug_protein_10xpre,
                             color=Treatment,
                             group=Treatment)) +
                         scale_colour_manual(breaks=c("Ambient", "Heated"), 
                                                             values=c("forestgreen","orange")) +
                         geom_point(aes(colour = Treatment), 
                                        position = position_dodge2(width = 0.4)) + 
                         stat_summary(fun.y="mean", size = 0.8, color = "black",
                                      position = position_dodge2(width = 0.4)) +
                         stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)), 
                                      fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                      geom = 'errorbar', width = 0.25, size = 1,
                                      position = position_dodge2(width = 10.4)) +
                         labs(title="PEPCK Activity (nM Pyruvate per ng protein)", 
                              x ="Temperature", 
                              y = expression("MDA"~(~uM^{-1}*~ug~protein^{-1}))) + #"CS activity per g protein") +
                        # scale_x_discrete(labels=c("22", "26", "30" , "34", "36")) +
                         # scale_y_continuous(expand = c(0, 0), limits = c(0, 0.00035), 
                                            # breaks = seq(0, 0.00035, by = 0.00007)) +
                         theme_classic() +
                         theme(panel.grid.major = element_blank(),  
                               panel.grid.minor = element_blank(), 
                               axis.title =element_text(size=12),
                               axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                               axis.text=element_text(size=10),
                               plot.title = element_text(size=12),
                               legend.position="none") + 
                        facet_wrap(~Treatment, scales= "free_x")

# plot all data above 0 - omit outliers and present only the heated treatment
Plot_MeanSE_ambient <- Master_Table %>% 
                  dplyr::filter(!(Treatment %in% 'Heated'))%>%
                  ggplot(aes(x = as.factor(Temperature),
                             y = MDA_uM_ug_protein,
                             color=Treatment)) +
                         scale_colour_manual(breaks=("Ambient"),
                                             values=("forestgreen")) +
                         geom_point(aes(colour = Treatment),
                                        position = position_dodge2(width = 0.4)) +
                         stat_summary(fun.y="mean", size = 0.8, color = "black",
                                      position = position_dodge2(width = 0.4)) +
                         stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)),
                                      fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                      geom = 'errorbar', width = 0.25, size = 1,
                                      position = position_dodge2(width = 10.4)) +
                         labs(title="uM MDA ug total protein", 
                              x ="Temperature",
                              y = expression("MDA"~(~uM^{-1}*~ug~protein^{-1}))) + #"CS activity per g protein") +
                        # scale_x_discrete(labels=c("22", "26", "30" , "34", "36")) +
                         # scale_y_continuous(expand = c(0, 0), limits = c(0, 0.00035),
                         #                    breaks = seq(0, 0.00035, by = 0.00007)) +
                         theme_classic() +
                         theme(panel.grid.major = element_blank(),
                               panel.grid.minor = element_blank(),
                               axis.title =element_text(size=12),
                               axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                               axis.text=element_text(size=10),
                               plot.title = element_text(size=12),
                               legend.position="none")

# Plot_MeanSE_ambient

# plot all data above 0 - omit outliers and present only the heated treatment
Plot_MeanSE_heated <- Master_Table %>% 
                  dplyr::filter(!(Treatment %in% 'Ambient')) %>%
                  ggplot(aes(x = as.factor(Temperature),
                             y = MDA_uM_ug_protein,
                             color=Treatment)) +
                         scale_colour_manual(breaks=("Heated"),
                                             values=("orange")) +
                         geom_point(aes(colour = Treatment),
                                        position = position_dodge2(width = 0.4)) +
                         stat_summary(fun.y="mean", size = 0.8, color = "black",
                                      position = position_dodge2(width = 0.4)) +
                         stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)),
                                      fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                      geom = 'errorbar', width = 0.25, size = 1,
                                      position = position_dodge2(width = 10.4)) +
                         labs(title="uM MDA ug total protein", 
                              x ="Temperature",
                              y = expression("MDA"~(~uM^{-1}*~ug~protein^{-1}))) + #"CS activity per g protein") +
                        # scale_x_discrete(labels=c("22", "26", "30" , "34", "36")) +
                         # scale_y_continuous(expand = c(0, 0), limits = c(0, 0.00035),
                         #                    breaks = seq(0, 0.00035, by = 0.00007)) +
                         theme_classic() +
                         theme(panel.grid.major = element_blank(),
                               panel.grid.minor = element_blank(),
                               axis.title =element_text(size=12),
                               axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                               axis.text=element_text(size=10),
                               plot.title = element_text(size=12),
                               legend.position="none")


ggarrange(Plot_MeanSE_ambient, Plot_MeanSE_heated)

# output 
pdf(paste("Output/Colorimetric_assays/PEPCK/PEPCK_AmbinetvHeated.pdf", sep =''), 
    width=10, 
    height=4)
print(Plot_MeanSE)
dev.off()


# pdf(paste("Output/Colorimetric_assays/Citrate_Synthase/Citrate_Synthase_Heated.pdf", sep =''), 
#     width=5, 
#     height=4)
# print(Plot_MeanSE_heated)
# dev.off()
# 
# pdf(paste("Output/Colorimetric_assays/Citrate_Synthase/Citrate_Synthase_Ambient.pdf", sep =''), 
#     width=5, 
#     height=4)
# print(Plot_MeanSE_ambient)
# dev.off()

```


# STATISTICS


```{r run stats}
library(rstatix)
Master <- rbind(Plate_1_CS_per.g.protein, Plate_2_CS_per.g.protein) %>% 
                  dplyr::filter(!(meanCs_Activty < 0  | meanCs_Activty > 0.0006)) %>%
                  dplyr::mutate(Temperature = case_when(
                    Temperature.Number == 1 ~ "22",
                    Temperature.Number == 2 ~ "26",
                    Temperature.Number == 3 ~ "30",
                    Temperature.Number == 4 ~ "34",
                    Temperature.Number == 5 ~ "38",
                  ))


Master %>% 
            group_by(Temperature) %>%
            anova_test(dv = meanCs_Activty, between = Treatment) %>%
            get_anova_table() %>%
            adjust_pvalue(method = "bonferroni")
  
```

