---
title: "Run_stats.Rmd"
author: "Sam Gurr"
date: "2024-07-03"
output: html_document
---

-   Last updates: July 3, 2024

# Thermal Performacne (pilot test) & INFISH projet with Indya Limpkin

## Used citrate synthase kit from Sigma aldrich on C virginica gill tissue homogenate, corrected to

total protein using pierce BCA protein assay (albumin standard)

## Load Libraries

### SET UP

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
# knitr::opts_knit$set(root.dir = "C:/Users/katherine.mcfarland/Documents/GitHub/EAD-ASEB-Cvirginica_Thermal_Performance/larvae") # Katie's
#knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Cvirginica_Thermal_Performance/RAnalysis") # Sam's
knitr::opts_knit$set(root.dir = "C:/Users/samuel.gurr/Documents/Github_repositories/Cvirginica_Thermal_Performance/RAnalysis") # Sam's work
```


## Load packages 

```{r setup, include=TRUE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(rstatix)
library(lmtest) # to receive p value from betareg model
# library(FSA) # for the Dun test post hoc for SRH non-parametric 2 way anova]
library(emmeans)
```


## load master data files 

```{r load calcualted data}

# citrate synthase
CS.df <- read.csv("Output/Colorimetric_Assays/Citrate_Synthase/Calc_Master_CSActivity.csv", sep = ',', head = T) %>% select(!X)

# PEPCK - stil in progress need calculated master file
PEPCK.df <- read.csv("Output/Colorimetric_assays/PEPCK/Calc_Master_PEPCK.csv", sep = ',', head = T) %>% select(!X)

# Succinate 
Succinate.df <- read.csv("Output/Colorimetric_Assays/Succinate/Calc_Master_Succinate.csv", sep = ',', head = T) %>% select(!X)

# Lipid peroxidation
LP.df <- read.csv("Output/Colorimetric_Assays/Lipid_Peroxidation/Calc_Master_Lipid_Peroxidation.csv", sep = ',', head = T) %>% select(!X)

# TAOC - a lot of missing data, did not output the calcualted master file

```

# STATISTICS

* Two approached as the following: 

  - (1) Are there differences between temperature X and the control at each condition during the ramp?
  To test this we run anovas at each timepoint pinning the challenge vs. the control  using 'rstatix; and Bonferroni correction
  
  - (2) Are there differences between temperatures within the challenge and timepoint within the control?
  This tests the challenge and control separately to determine whether we recorded a response to temperature (in the challange) or 
  another factor (handling, circadian rhythm, etc.) elicited differences over time in the control (static temperature) 


```{r run stats}
library(rstatix)

# citrate synthase :::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::

# (1) Each timepoint temperature challenge vs control
CS.stats <-
CS.df %>%   
  group_by(Timepoint) %>%
  anova_test(dv = meanCs_Activity_ug_protein, between = Treatment) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")

# (2) within treatment, effect of temperature / time
CS.stats_heated <-
CS.df %>% 
  filter(Treatment %in% 'Heated') %>% 
  anova_test(dv = meanCs_Activity_ug_protein, between = Temperature) %>% 
  get_anova_table()

CS.stats_ambient <-
CS.df %>% 
  filter(Treatment %in% 'Ambient') %>% 
  anova_test(dv = meanCs_Activity_ug_protein, between = Timepoint) %>% 
  get_anova_table()


write.csv(CS.stats, file = "Output/Colorimetric_assays/Citrate_Synthase/ANOVA_each_timepoint_CS.csv") # write csv
write.csv(CS.stats_heated, file = "Output/Colorimetric_assays/Citrate_Synthase/ANOVA_heated_CS.csv") # write csv
write.csv(CS.stats_ambient, file = "Output/Colorimetric_assays/Citrate_Synthase/ANOVA_ambient_CS.csv") # write csv

# Succinate  :::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::

# (1) Each timepointtemperature challange vs control
Succinate.stats <-
Succinate.df %>%   
  group_by(Timepoint) %>%
  anova_test(dv = meanSuccinate_per_ug_protein_5xpre, between = Treatment) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")

Succinate.df %>%   
  group_by(Timepoint) %>%
  anova_test(dv = meanSuccinate_per_ug_protein_10xpre, between = Treatment) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")

Succinate.df %>%   
  group_by(Timepoint) %>%
  anova_test(dv = meanSuccinate_per_ug_protein_10xpost, between = Treatment) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")

Succinate.df %>%   
  group_by(Timepoint) %>%
  anova_test(dv = meanSuccinate_per_ug_protein_10xpost30, between = Treatment) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")

# (2) within treatment, effect of temperature / time
Succinate.stats_heated <-
Succinate.df %>% 
  filter(Treatment %in% 'Heated') %>% 
  mutate(Temperature = as.factor(Temperature)) %>% 
  anova_test(dv = meanSuccinate_per_ug_protein_5xpre, between = Temperature) %>% 
  get_anova_table()

Succinate.stats_ambient <-
Succinate.df %>% 
  filter(Treatment %in% 'Ambient') %>% 
  mutate(Timepoint = as.factor(Timepoint)) %>% 
  anova_test(dv = meanSuccinate_per_ug_protein_5xpre, between = Timepoint) %>% 
  get_anova_table()


mod <- aov(meanSuccinate_per_ug_protein_5xpre ~ Treatment*Temperature, 
           data = 
             (Succinate.df %>% 
                select(meanSuccinate_per_ug_protein_5xpre,Treatment,Temperature) %>% 
                na.omit())
              )
summary(mod)


write.csv(Succinate.stats, file = "Output/Colorimetric_assays/Succinate/ANOVA_Succinate.csv") # write csv
write.csv(Succinate.stats_heated, file = "Output/Colorimetric_assays/Succinate/ANOVA_heated_CS.csv") # write csv
write.csv(Succinate.stats_ambient, file = "Output/Colorimetric_assays/Succinate/ANOVA_ambient_CS.csv") # write csv


# Lipid peroxidation

# (1) Each timepointtemperature challange vs control
LP.df %>%   
  group_by(Timepoint, Treatment) %>% 
  summarise(n = n()) # timepoint 4 does not have Ambient, wont run below
  

LP.stats <-
LP.df %>%   
  filter(!Timepoint %in% 4) %>% # as shown abovem we need to omit for this to run
  group_by(Timepoint) %>%
  anova_test(dv = MDA_uM_ug_protein, between = Treatment) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")

# (2) within treatment, effect of temperature / time
LP.stats_heated <-
LP.df %>% 
  filter(Treatment %in% 'Heated') %>% 
  mutate(Temperature = as.factor(Temperature)) %>% 
  anova_test(dv = MDA_uM_ug_protein, between = Temperature) %>% 
  get_anova_table()

LP.stats_ambient <-
LP.df %>% 
  filter(Treatment %in% 'Ambient') %>% 
  mutate(Timepoint = as.factor(Timepoint)) %>% 
  anova_test(dv = MDA_uM_ug_protein, between = Timepoint) %>% 
  get_anova_table()


write.csv(LP.stats, file = "Output/Colorimetric_assays/Lipid_Peroxidation/ANOVA_Lipid_Peroxidation.csv") # write csv
write.csv(LP.stats_heated, file = "Output/Colorimetric_assays/Lipid_Peroxidation/ANOVA_heated_CS.csv") # write csv
write.csv(LP.stats_ambient, file = "Output/Colorimetric_assays/Lipid_Peroxidation/ANOVA_ambient_CS.csv") # write csv

# PEPCK 

# (1) Each timepointtemperature challange vs control
PEPCK.df %>%   
  group_by(Timepoint, Treatment) %>% 
  summarise(n = n()) # timepoint 3 does not have Ambient, wont run below
  

PEPCK.stats <-
  PEPCK.df %>%   
  filter(!(Timepoint %in% 3 | PEPCK_Activity_nmolPyruvante_ngProtein > 1.7)) %>% # as shown above we need to omit for this to run
  # filter(!Volume %in% 10 ) %>% 
  filter(PEPCK_Activity_nmolPyruvante_ngProtein > 0) %>% 
  group_by(Timepoint) %>%
  anova_test(dv = PEPCK_Activity_nmolPyruvante_ngProtein, between = Treatment) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")


# (2) within treatment, effect of temperature / time
PEPCK.stats_heated <-
PEPCK.df %>% 
  filter(!(Timepoint %in% 3 | PEPCK_Activity_nmolPyruvante_ngProtein > 1.7)) %>% # as shown above we need to omit for this to run
  filter(Treatment %in% 'Heated') %>% 
  mutate(Temperature = as.factor(Temperature)) %>% 
  anova_test(dv = PEPCK_Activity_nmolPyruvante_ngProtein, between = Temperature) %>% 
  get_anova_table()

PEPCK.stats_ambient <-
PEPCK.df %>% 
  filter(!(Timepoint %in% 3 | PEPCK_Activity_nmolPyruvante_ngProtein > 1.7)) %>% # as shown above we need to omit for this to run
  filter(Treatment %in% 'Ambient') %>% 
  mutate(Timepoint = as.factor(Timepoint)) %>% 
  anova_test(dv = PEPCK_Activity_nmolPyruvante_ngProtein, between = Timepoint) %>% 
  get_anova_table()



write.csv(PEPCK.stats, file = "Output/Colorimetric_assays/PEPCK/ANOVA_PEPCK.csv") # write csv
write.csv(PEPCK.stats_heated, file = "Output/Colorimetric_assays/PEPCK/ANOVA_heated_CS.csv") # write csv
write.csv(PEPCK.stats_ambient, file = "Output/Colorimetric_assays/PEPCK/ANOVA_ambient_CS.csv") # write csv

```

