---
title: "Total Antioxidant Capacity.Rmd"
author: "Sam Gurr"
date: "2024-07-23"
output: html_document
---

-   Last updates: July 23, 2024

# Thermal Performacne (pilot test) & INFISH projet with Indya Limpkin

## Used Total Antioxidant capacity kit from Cell Biolabs using C virginica gill tissue homogenate, corrected to

total protein using pierce BCA protein assay (albumin standard)

## Load Libraries

### SET UP

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
# knitr::opts_knit$set(root.dir = "C:/Users/katherine.mcfarland/Documents/GitHub/EAD-ASEB-Cvirginica_Thermal_Performance/larvae") # Katie's
knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/EAD-ASEB-Cvirginica_Thermal_Performance/RAnalysis") # Sam's
#knitr::opts_knit$set(root.dir = "C:/Users/samuel.gurr/Documents/Github_repositories/EAD-ASEB-Cvirginica_Thermal_Performance/RAnalysis") # Sam's work
```


## Load packages 

```{r setup, include=TRUE}

# library(lmtest) # to receive p value from betareg model
# library(FSA) # for the Dun test post hoc for SRH non-parametric 2 way anova]
# library(emmeans)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(car)
# library(lmerTest)
library(tidyr)
# library(reshape2)
library(ggpubr)
# library(nlme)
# library(rcompanion) # to run the Schrier -Ray-Hare non parametric 2 way 
library(ggpmisc) # stat_poly for inserting equation and R2 for ggplot line 
```

## Function to load files

-   At your service, from 'spec to table' ...I'm so funny

```{r Spec_to_table}

Spec_to_table <- function(datapath, platenumber, daterecorded) {
  
# import the data in raw text form exported from the spectrophotometer
        rawtextfile <- read.delim(datapath,
                              header = F,
                              skip=3,
                              fileEncoding="latin1")
  

        rawmatrix   <- as.matrix(rawtextfile[c(1:8),c(3:14)])
        colnames(rawmatrix) = c("1","2","3","4","5","6","7","8","9","10","11","12")
        rownames(rawmatrix) = c("A","B","C","D","E","F","G","H")
        
        raw_table <- as.data.frame.table(rawmatrix, responseName = "value") %>% 
                                    dplyr::rename(well_row=Var1, well_column=Var2, Abs_412nm=value) %>% 
                                    dplyr::mutate(well=paste0(well_row,well_column),
                                                  Run_date =daterecorded,
                                                  Plate=platenumber) %>% 
                                    dplyr::select(-c(well_row,well_column))
        # reorder the columns because I prefer it this way..
        
        raw_table_ordered <- raw_table[,c(3,4,2,1)]
}


```

-   Use the spec_to_table to import raw spec txt files and compile to a table, yes that is what it does

-   add a column for the time, what is the time? - timepoints 1, 2, 3, 4 etc. represent 5 minute intervals starting at time 0 Example:

    -   timepoint1 = 0 minutes
    -   timepoint 2 = 5 minutes
    -   and so on...

```{r load TAOC raw data and build master files}

# plate 1 
TAOC_Plate1_before <- Spec_to_table('Data/Total_Antioxidant_Capacity/20240709_TAOC_Plate1_beforeCu.txt', 1, 20240709) %>% dplyr::mutate(Time = 'BEFORE Cu activation', Plate = 1) %>% dplyr::rename(Abs_490nm = Abs_412nm)
TAOC_Plate1_after  <- Spec_to_table('Data/Total_Antioxidant_Capacity/20240709_TAOC_Plate1_afterCu.txt', 1, 20240709) %>% dplyr::mutate(Time = 'AFTER Cu activation', Plate = 1) %>% dplyr::rename(Abs_490nm = Abs_412nm)

# plate 2
TAOC_Plate2_before <- Spec_to_table('Data/Total_Antioxidant_Capacity/20240709_TAOC_Plate2_beforeCu.txt', 2, 20240709) %>% dplyr::mutate(Time = 'BEFORE Cu activation', Plate = 2) %>% dplyr::rename(Abs_490nm = Abs_412nm)
TAOC_Plate2_after  <- Spec_to_table('Data/Total_Antioxidant_Capacity/20240709_TAOC_Plate2_afterCu.txt', 2, 20240709) %>% dplyr::mutate(Time = 'AFTER Cu activation', Plate = 2) %>% dplyr::rename(Abs_490nm = Abs_412nm)

# # plate 1 & 2 master files
TAOC_Plate1_master <- do.call("rbind", list(TAOC_Plate1_before, TAOC_Plate1_after))

TAOC_Plate2_master <- do.call("rbind", list(TAOC_Plate2_before, TAOC_Plate2_after))
```

```{r load Total protein correction data for citrate synthase}

# plate 1 
TP_Plate1_master <- Spec_to_table('Data/Total_Protein_Total_Antioxidant_Capacity/Total_Protein_TAOC_Plate1.txt', 1, 20240711)  %>%  dplyr::rename(Abs_562nm = Abs_412nm)

# plate 2
TP_Plate2_master <- Spec_to_table('Data/Total_Protein_Total_Antioxidant_Capacity/Total_Protein_TAOC_Plate2.txt', 2, 20240711)   %>%  dplyr::rename(Abs_562nm = Abs_412nm)

```


```{r load reference metadata and call target datasets}

metadata <- read.csv('Data/TAOC_metadata.csv', sep = ",", header=T)

# rename a few columns so they can merge with the data (in next chunk)

# Plate 1 
metadata.TAOC_Plate_1 <- metadata %>% 
                        dplyr::filter(Assay.Type %in% "Total Antioxidant Capacity Assay " &
                                        Plate.Number == 1) %>% 
                        dplyr::rename(Plate = Plate.Number,
                                      well  = Well.ID)
# Plate 2
metadata.TAOC_Plate_2 <- metadata %>% 
                        dplyr::filter(Assay.Type %in% "Total Antioxidant Capacity Assay " & 
                                        Plate.Number == 2) %>% 
                        dplyr::rename(Plate = Plate.Number,
                                      well  = Well.ID)



# Plate 1 
metadata.TP_Plate_1 <- metadata %>% 
                        dplyr::filter(Assay.Type %in% "Total Protein" & 
                                        Plate.Number == 1) %>% 
                        dplyr::rename(Plate = Plate.Number,
                                      well  = Well.ID)
# Plate 2
metadata.TP_Plate_2 <- metadata %>% 
                        dplyr::filter(Assay.Type %in% "Total Protein" & 
                                        Plate.Number == 2) %>% 
                        dplyr::rename(Plate = Plate.Number,
                                      well  = Well.ID)



```


* merge metadata with data 
```{r merge metadata with data}

#  TAOC
# Plate 1 
Plate1 <- merge(metadata.TAOC_Plate_1, TAOC_Plate1_master, by = c('Plate', 'well'))                          
nrow(TAOC_Plate1_master) == nrow(Plate1) # must be TRUE
nrow(Plate1) # 192 rows
Plate1 <- subset(Plate1, !is.na(Sample.Type)) # now 574  total rows with NAs omitted (not sample, blank nor stnadard)
nrow(Plate1) # 192 rows - no rows omitted, full plate


# Plate 2
Plate2 <- merge(metadata.TAOC_Plate_2, TAOC_Plate2_master, by = c('Plate', 'well'))                          
nrow(TAOC_Plate2_master) == nrow(Plate2) # must be TRUE
nrow(Plate2) # 192 rows
Plate2 <- subset(Plate2, !is.na(Sample.Type)) # now 328  total rows with NAs omitted (not sample, blank nor stnadard)
nrow(Plate2) # 168 rows, several rows of the plate were NA


# Total protein
# Plate 1 
Plate1.TP <- merge(metadata.TP_Plate_1, TP_Plate1_master, by = c('Plate', 'well'))                          
nrow(TP_Plate1_master) == nrow(Plate1.TP) # must be TRUE
nrow(Plate1.TP) # 96 
Plate1.TP <- subset(Plate1.TP, !is.na(Sample.Type)) # now 87 total rows with NAs omitted (not sample, blank nor stnadard)
nrow(Plate1.TP) # 95 - one NA


# Plate 2 
Plate2.TP <- merge(metadata.TP_Plate_2, TP_Plate2_master, by = c('Plate', 'well'))                          
nrow(TP_Plate2_master) == nrow(Plate2.TP) # must be TRUE
nrow(Plate2.TP) # 96
Plate2.TP <- subset(Plate2.TP, !is.na(Sample.Type)) # now 68 total rows with NAs omitted (not sample, blank nor stnadard)
nrow(Plate2.TP) # 80 - several NAs not a full plate

```


* subset standards, blanks, and samples
```{r subset data types}

# TAOC
# Plate 1 
Plate1.samples   <- Plate1 %>% dplyr::filter(Sample.Type %in% 'Sample')
Plate1.blanks    <- Plate1 %>% dplyr::filter(Sample.Type %in% c('Blank', 'Blank ')) # space, why... whatevs
Plate1.standards <- Plate1 %>% dplyr::filter(Sample.Type %in% 'Standard')

(nrow(Plate1.samples) + 
    nrow(Plate1.blanks) + 
      nrow(Plate1.standards)) == nrow(Plate1) # sanity check, must be true


# Plate 2
Plate2.samples   <- Plate2 %>% dplyr::filter(Sample.Type %in% 'Sample')
Plate2.blanks    <- Plate2 %>% dplyr::filter(Sample.Type %in%  c('Blank', 'Blank ')) # space, why... whatevs
Plate2.standards <- Plate2 %>% dplyr::filter(Sample.Type %in% 'Standard')

(nrow(Plate2.samples) + 
    nrow(Plate2.blanks) + 
    nrow(Plate2.standards)) == nrow(Plate2) # sanity check, must be true






# Total protein
# Plate 1 
Plate1_TP.samples    <- Plate1.TP %>% dplyr::filter(Sample.Type %in% 'Sample')
# Plate1_TP.blanks     <- Plate1.TP %>% dplyr::filter(Sample.Type %in% 'Blank')
Plate1_TP.standards  <- Plate1.TP %>% dplyr::filter(Sample.Type %in% c('Standard', 'Blank'))

(nrow(Plate1_TP.samples) + 
    # nrow(Plate1_TP.blanks) + 
      nrow(Plate1_TP.standards)) == nrow(Plate1.TP) # sanity check, must be true


# Plate 2
Plate2_TP.samples    <- Plate2.TP %>% dplyr::filter(Sample.Type %in% 'Sample')
# Plate2_TP.blanks     <- Plate2.TP %>% dplyr::filter(Sample.Type %in% 'Blank') 
Plate2_TP.standards  <- Plate2.TP %>% dplyr::filter(Sample.Type %in% c('Standard', 'Blank'))

(nrow(Plate2_TP.samples) + 
    # nrow(Plate2_TP.blanks) + 
      nrow(Plate2_TP.standards)) == nrow(Plate2.TP) # sanity check, must be true

```

## Total protein calculations 

```{r total protein standard curve}

# assign known BCA values to standard IDs A-I
BCA_standards <- rbind(Plate1_TP.standards, Plate2_TP.standards) %>%  
  dplyr::mutate(BCA_ug_mL = case_when(Standards %in% 'A' ~ 2000,
                                      Standards %in% 'B' ~ 1500,
                                      Standards %in% 'C' ~ 1000,
                                      Standards %in% 'D' ~ 750,
                                      Standards %in% 'E' ~ 500,
                                      Standards %in% 'F' ~ 250,
                                      Standards %in% 'G' ~ 125,
                                      Standards %in% 'H' ~ 25,
                                      Standards %in% 'I' ~ 0)) %>% 
  dplyr::select(Plate, Standards, BCA_ug_mL, Abs_562nm)

# Run standard curve, calculate totalprotein 
BCA_background_zero <- BCA_standards %>% 
                        dplyr::filter(Standards %in% 'I') %>% # the zero standard
                        dplyr::group_by(Plate, Standards, BCA_ug_mL) %>% # group by to get the means
                        dplyr::summarise_each(funs(mean,sd,se=sd(.)/sqrt(n()))) # get all the stats 


# Plate 1, blank to correct by is 0.11120
# Plate 2, blank to correct by is 0.12215



# Absorbance corrected - take the mean of any duplicates
BCA_standards_means <- BCA_standards %>% 
                        #dplyr::filter(!Type %in% 'Standard_0') %>% 
                        dplyr::mutate(Abs_562nm_cor = 
                                     case_when(Plate == 1 ~ (Abs_562nm-0.11120),
                                               Plate == 2 ~ (Abs_562nm-0.12215) ) ) %>% 
                        dplyr::select(-Abs_562nm) %>% 
                        dplyr::group_by(Plate, Standards, BCA_ug_mL) %>% # group by to get the means
                        dplyr::summarise_each(funs(mean,sd,se=sd(.)/sqrt(n()))) # get all the stats 


# plot it insert the quadratic formaula using ggpmisc
BCA_stand_plots_quadratic <- BCA_standards_means %>% 
                    # QUADRATIC SMOOTH LINE WORKS BEST HERE (MANUFACTURERS INSTRUCTIONS)
                     ggplot(aes(y=mean, x=BCA_ug_mL)) + 
                        geom_point() +
                        theme_bw() +
                        labs(y= "Net Abs 562nm", x = "Protein Concentration in ug/mL") +
                        #geom_line() +
                        #stat_poly_line(color='red') +
                        #geom_smooth() +
                        stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1) +
                        stat_poly_eq(parse=T, aes(label = ..eq.label..), formula=y ~ x + I(x^2)) +
                        ggtitle('Total protein: Quadratic curve') +
                        #stat_poly_eq(use_label(c("eq", "R2"))) +
                        facet_wrap(~Plate) 

# output 
pdf(paste("Output/Colorimetric_assays/Total_Antioxidant_Capacity/Standard_Curve_BCA_TAOCcorrection.pdf", sep =''), 
    width=10, 
    height=7)
print(ggarrange(BCA_stand_plots_quadratic))
dev.off()

```



* Use BCA standard curves to calculate TP 

```{r calc TP}

# Standard curve, Plate 1 equation y = 0.0154 + 0.00125x - 1.17x10^-7x^2 - need to solve for x!
# Standard curve, Plate 2 equation y = 0.0201 + 0.00133x - 1.32x10^-7x^2 - need to solve for x!

# Standard curve, Plate 1
a1 <- -1.17*10^-7
b1 <- 0.00125
c1 <- 0.0154
# EQ: (-(b1) + sqrt( (b1^2) - (4*(((a1)-Abs_562nm_cor))*(c1)) ))/(2*a1)

# Standard curve, Plate 2
a2 <- -1.32*10^-7
b2 <- 0.00133
c2 <- 0.0201
# EQ: (-(b2) + sqrt( (b2^2) - (4*a2*(c2-Abs_562nm_cor)) ) ) / (2*a2)


# linear equation plate 1 == (Abs_562nm_cor - 0.192)/0.000993
# linear equation plate 2 == (Abs_562nm_cor - 0.224)/0.000911


# IMPORTANT! we used 25 ul of the standards and 25 ul of the unknowns (samples) 
# therefore we can interpret the unknown direct to the the standard curve without having 
# to account for addition factors, fot example, if we used 5 ul unknown (sample) we would have to adjust 
# by multiplying by 5 to reach the standard curve 

V = 0.025 # 25 ul or 0.025 mL

# Plate 1, blank to correct by is 0.11120
# Plate 2, blank to correct by is 0.12215

# Sanity check Lets look at the absorbance vs. totla protein concentration data 

TotalProtein_final <- rbind(Plate1_TP.samples, Plate2_TP.samples) %>% 
                      dplyr::select(Plate,Tank.Number,Temperature.Number,Abs_562nm) %>% 
                      dplyr::filter(!Abs_562nm > 3.99) %>% # data as 4.00 is above the detection limit, omit
                      dplyr::mutate(Unique_ID = 
                                      paste0('Plate',Plate,'_',
                                             'Tank',Tank.Number, '_',
                                             'Temperature', Temperature.Number)) %>% # unique ID t0 group by
                      dplyr::mutate(Abs_562nm_cor = # correct the raw abs, subtract background
                                   case_when(Plate == 1 ~ (Abs_562nm-0.11120), # for plate 1
                                             Plate == 2 ~ (Abs_562nm-0.12215) ) ) %>% # for plate 2 
                      dplyr::mutate(TotalProtein_ug_mL = 
                                    case_when(
                                      # linear fr neg discrim. - luckily only two values from plate 2
                                      # Scallop_ID %in% c(33, 51) ~ 
                                      #   ((Abs_562nm_cor - 0.224)/0.000911),
                                      # quadratic for Plate 1
                                      Plate == 1 ~ 
                                        ((-(b1) + sqrt( (b1^2) - (4*a1*(c1-Abs_562nm_cor)) ) ) / (2*a1)), 
                                      # quadratic for plate 2
                                      Plate == 2 ~ 
                                        ((-(b2) + sqrt( (b2^2) - (4*a2*(c2-Abs_562nm_cor)) ) ) / (2*a2)) ),
                                    # ug per mL concentration to ug in 25 ul sample 
                                    TotalProtein_ug = TotalProtein_ug_mL*V)

# View(TotalProtein_final)
nrow(TotalProtein_final) # 137

# NOTE! I found that some of my discriminants are negative (b^2 - 4ac) IN PLATE 2!! (as 'NaN)
# so Im going to extrapolate from a linear regression curve based on the final samples 

# first lets plot absorbance and actual concentration calacuated 
calc_BCA_plot <- TotalProtein_final %>% 
                    ggplot(aes(y = Abs_562nm_cor, 
                               x  = TotalProtein_ug_mL)) +
                          geom_point() + 
                    theme_bw() + 
                    ggtitle('Total protein: Calculated BCA by Net Absorbance') +
                    facet_wrap(~Plate)
calc_BCA_plot 


#print
pdf(paste("Output/Colorimetric_assays/Total_Antioxidant_Capacity/Standard_Curve_BCA_TAOCcorrection.pdf", sep =''), 
    width=10, 
    height=14)
ggarrange(BCA_stand_plots_quadratic,
          calc_BCA_plot,
          nrow = 2)
dev.off()
# write csv
write.csv(TotalProtein_final, file = "Output/Colorimetric_assays/Total_Antioxidant_Capacity/Calc_Master_BCA_TAOCcorrection.csv")


```

## Total Antioxidant Capacity calculations

### Step 1: Net Absorbance for standards and samples: Calculate by simply subtracting the 
final reading from the inital, this is the 'net absorbance'

Files: 
* Plate1.samples, Plate1.standards
* Plate2.samples, Plate2.standards
 
### Step 2: Generate a Standard Curve - plot the net absorbance of standards against their known uric acid concentration

### Step 3a: Calculate mM uric acid equivalents (UAE) - extrapolate the the ruic acid
concentration from the OD 490 nm value on uric acid standard curve (output of Step 2) 

### Step 3b: Calculate uM Copper Reducing Equivalents (CRE) - simply muliply UAE * 2189 (mM Cu++)
the CRE value is proportional to the sample's Total antioxidant capacity 

### Step 4: Correct to individual's per gram total protein - NOTE: make sure you correct using the corresponding 
volume of fill homogenate for each assay, Pierce BCA assay used 25 ul of homogenate whereas the Oxiselect 
assay (TAOC) used only 20 ul of gill homogenate. 

* file need: TotalProtein_final

### Plot data and run statistical models for treatment effects
 * you can  use the ggplot call here to generate y axis units labs(y=expression("Total Antioxidant Capacity"~(~?M~CRE[red]^{-1}*g~total~protein^{-1}))


* STEP 1: Net absorbance

  - final [AFTER Cu activation] - initial [BEFORE Cu activation]

```{r Step 1 TAOC - Net Absorbance}

# plate 1
Plate1_samples_NetAbs <- Plate1.samples %>%  dplyr::select(!c(Standards, Run_date)) %>% 
                            tidyr::pivot_wider(names_from = Time, values_from = Abs_490nm) %>% 
                            dplyr::mutate(NetAbs_OD490nm = `AFTER Cu activation` - `BEFORE Cu activation`)

Plate1_standards_NetAbs <- Plate1.standards %>%  dplyr::select(!c(Run_date, Temperature.Number, Tank.Number, Volume)) %>% 
                            tidyr::pivot_wider(names_from = Time, values_from = Abs_490nm) %>% 
                            dplyr::mutate(NetAbs_OD490nm = `AFTER Cu activation` - `BEFORE Cu activation`)

# plate 2
Plate2_samples_NetAbs <- Plate2.samples %>%  dplyr::select(!c(Standards, Run_date)) %>% 
                            tidyr::pivot_wider(names_from = Time, values_from = Abs_490nm) %>% 
                            dplyr::mutate(NetAbs_OD490nm = `AFTER Cu activation` - `BEFORE Cu activation`)

Plate2_standards_NetAbs <- Plate2.standards %>%  dplyr::select(!c(Run_date, Temperature.Number, Tank.Number, Volume)) %>% 
                            tidyr::pivot_wider(names_from = Time, values_from = Abs_490nm) %>% 
                            dplyr::mutate(NetAbs_OD490nm = `AFTER Cu activation` - `BEFORE Cu activation`)

```


* STEP 2: Generate standard curve

  - plot the net absorbance of standards against their known uric acid concentration
  
  - NOTE: the first plate has the only good TAOC standard curve, apply this formula to both plates
  
```{r Step 2 TAOC - Standard Curve}

# Assign known Uric acid concentration in nM for each standard
Uric_Acid_standards <- rbind(Plate1_standards_NetAbs, Plate2_standards_NetAbs) %>%  
                        dplyr::mutate(UricAcid_mM = case_when(Standards == 1 ~ 1,
                                                            Standards == 2 ~ 0.5,
                                                            Standards == 3 ~ 0.25,
                                                            Standards == 4 ~ 0.125,
                                                            Standards == 5 ~ 0.0625,
                                                            Standards == 6 ~ 0.03125,
                                                            Standards == 7 ~ 0.0156,
                                                            Standards == 8 ~ 0.0078,
                                                            Standards == 9 ~ 0.0039,
                                                            Standards == 10 ~ 0)) %>% 
                        dplyr::select(Plate, Standards, UricAcid_mM, NetAbs_OD490nm)



# Average by duplicate IDs (Plate and Stnadard ID)
# note that mean se and sd are of the net absorbance values
Uric_Acid_standards_means <- Uric_Acid_standards %>% 
                        dplyr::group_by(Plate, Standards, UricAcid_mM) %>% # group by IDs
                        dplyr::summarise_each(funs(mean,sd,se=sd(.)/sqrt(n()))) # get all the stats 


# plot it insert the linear formaula using ggpmisc
Uric_Acid_standards_linear <- Uric_Acid_standards_means %>% 
                    # QUADRATIC SMOOTH LINE WORKS BEST HERE (MANUFACTURERS INSTRUCTIONS)
                     # dplyr::filter(!(Plate %in% 2 & Standards %in% 'D')) %>% # hash me out to test
                     ggplot(aes(y=mean, x=UricAcid_mM)) + 
                        geom_point() +
                        theme_bw() +
                        labs(y= "Net Abs 480nm", x = "Uric Acid concentration mM") +
                        #geom_line() +
                        #stat_poly_line(color='red') +
                        #geom_smooth() +
                        stat_smooth(method = "lm", formula = y ~ x, size = 1) +
                        stat_poly_eq(parse=T, aes(label = ..eq.label..), formula=y ~ x) +
                        ggtitle('Uric Acid Standards: Quadratic curve') +
                        #stat_poly_eq(use_label(c("eq", "R2"))) +
                        facet_wrap(~Plate) 

# output 
pdf(paste("Output/Colorimetric_assays/Total_Antioxidant_Capacity/Standard_Curve_UricAcid.pdf", sep =''), 
    width=10, 
    height=5)
print(ggarrange(Uric_Acid_standards_linear))
dev.off()

```



### STEP 3a: Calculate mM uric acid equivalents (UAE) 

* do this for sample file: Plate1_samples_NetAbs, Plate2_samples_NetAbs

  - standard curve formula: y = 0.0248 + 1.56x; y == Net absorbance at 490 nm; x = uric acid concentration
  
* extrapolate the the ruic acid concentration from the OD 490 nm value on uric acid standard curve (output of Step 2) 

```{r Step 3a TAOC - Calculate mM uric acid equivalents (UAE) }

# same formulate for both plates, plate 2 had an error so use the same curve from plate 1 (previous chunk)
# plate 1 
Plate1_samples_UAE <- Plate1_samples_NetAbs %>% 
                            dplyr::mutate(UAE = (NetAbs_OD490nm - 0.0248) / 1.56)

# plate 2
Plate2_samples_UAE <- Plate2_samples_NetAbs %>% 
                            dplyr::mutate(UAE = (NetAbs_OD490nm - 0.0248) / 1.56)
```

**NOTE** : Something is terribly wrong with plate 2, perhaps a reagent was not added, the net absorbance is very low 
for BOTH standards and samples, **we will need to omit this entire plate from our analysis**

### STEP 3b: Calculate uM Copper Reducing Equivalents (CRE)

* do this for sample file: Plate1_samples_UAE, Plate2_samples_UAE

  - simply muliply UAE * 2189 (mM Cu++), the CRE value is proportional to the sample's Total antioxidant capacity 

```{r Step 3b TAOC - Calculate uM Copper Reducing Equivalents (CRE)}

# plate 1 
Plate1_samples_CRE <- Plate1_samples_UAE %>% 
                            dplyr::mutate(CRE = UAE * 2189)

# plate 2
Plate2_samples_CRE <- Plate2_samples_UAE %>% 
                            dplyr::mutate(CRE = UAE * 2189)

```

### STEP 4: Correct to total protein

* **Note** Plate 2 is a wash, do not proceed with it, **just proceed with plate 1**

```{r Step 4 TAOC - Correct to individual total gram protein}

# Total protein means per sample (group by and summarise the duplicates!)
TotalProtein_Plate1_MEANS <- TotalProtein_final %>% 
                                dplyr::filter(Plate == 1) %>% 
                                dplyr::select(Plate, Tank.Number, Temperature.Number, TotalProtein_ug_mL) %>% 
                                dplyr::group_by(Plate, Tank.Number, Temperature.Number) %>% # group by IDs
                                dplyr::summarise_each(funs(mean,sd,se=sd(.)/sqrt(n()))) %>% 
                                dplyr::select(!c(sd, se)) %>% 
                                dplyr::rename(TotalProtein_ug_mL = mean)
# TotalProtein_ug was corrected to the 25 ul used, meaning it is in total protein per ul; we used 
# 20 ul in the TAOC assay so multiple by 20 for the CRE correction below!


# CRE means per sample (group by and summarise the duplicates!)
Plate1_samples_CRE_MEANS <- Plate1_samples_CRE %>% 
                                dplyr::select(Plate, Tank.Number, Temperature.Number, CRE) %>% 
                                dplyr::group_by(Plate, Tank.Number, Temperature.Number) %>% # group by IDs
                                dplyr::summarise_each(funs(mean,sd,se=sd(.)/sqrt(n()))) %>% 
                                dplyr::select(!c(sd, se)) %>% 
                                dplyr::rename(CRE = mean)

# CRE corrected to total protein per sample -  20 ul or 0.025 mL
Plate1_samples_CRE.final <- merge(Plate1_samples_CRE_MEANS, TotalProtein_Plate1_MEANS, 
                                  by = c('Plate', 'Tank.Number', 'Temperature.Number')) %>% 
                              dplyr::mutate(CRE_per_ugprotein = CRE / (TotalProtein_ug_mL*0.020),
                                            Treatment = case_when(Tank.Number %in% c(1:7) ~ 'Ambient',
                                                                  Tank.Number %in% c(8:14) ~ 'Heated'))

 
```

### Plot data and run statistical models for treatment effects


```{r Plot data}

# Plate 1 only - no plate 2

MEanSE_TAOC <-  Plate1_samples_CRE.final%>% 
                        ggplot(aes(x = Temperature.Number, 
                                   y = CRE_per_ugprotein, 
                                   colour=Treatment,
                                   group=Treatment)) +
                               scale_colour_manual(breaks=c("Ambient", "Heated"), 
                                                            values=c("forestgreen","orange")) +
                               geom_point(aes(colour = Treatment), 
                                              position = position_dodge2(width = 0.4)) + 
                               stat_summary(fun.y="mean", size = 0.8, color = "black",
                                            position = position_dodge2(width = 0.4)) +
                               stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)), 
                                            fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                            geom = 'errorbar', width = 0.25, size = 1,
                                            position = position_dodge2(width = 0.4)) +
                               labs(title="Total Antioxidant Capacity", 
                                    x ="Temperature",
                                    y=expression("Total Antioxidant Capacity"~(~uM~CRE[red]^{-1}*~ug~total~protein^{-1}))) +                              
                               # scale_x_discrete(labels=c("22", "26", "30" , "34", "38")) +
                               scale_y_continuous(expand = c(0, 0), 
                                                   limits = c(0, 8), 
                                                   breaks = seq(0, 8, 
                                                                by = 2)) +
                               theme_classic() +
                               theme(panel.grid.major = element_blank(),  
                                     panel.grid.minor = element_blank(), 
                                     # axis.title.y=element_blank(),
                                     axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                                     axis.text=element_text(size=8),
                                     plot.title = element_text(size=12),
                                     legend.position="none") 


# Ambient only 

MeanSE_TAOC_ambient <-  Plate1_samples_CRE.final %>% 
                          dplyr::filter(!(Treatment %in% 'Heated')) %>%
                          dplyr::mutate(Temperature = case_when(
                            Temperature.Number == 1 ~ "22_t1",
                            Temperature.Number == 2 ~ "22_t2",
                            Temperature.Number == 3 ~ "22_t3",
                            Temperature.Number == 4 ~ "22_t4",
                            Temperature.Number == 5 ~ "22_t5",
                          )) %>%                               
                               ggplot(aes(x = factor(Temperature), 
                                   y = CRE_per_ugprotein, 
                                   colour=Treatment,
                                   group=Treatment)) +
                               scale_colour_manual(breaks=("Ambient"), 
                                                            values=("forestgreen")) +
                               geom_point(aes(colour = Treatment), 
                                              position = position_dodge2(width = 0.4)) + 
                               stat_summary(fun.y="mean", size = 0.8, color = "black",
                                            position = position_dodge2(width = 0.4)) +
                               stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)), 
                                            fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                            geom = 'errorbar', width = 0.25, size = 1,
                                            position = position_dodge2(width = 0.4)) +
                               labs(title="Total antioxidant capacity: Ambient", 
                                    x ="Temperature",
                                    y=expression("TAOC"~(~uM~CRE[red]^{-1}*~ug~protein^{-1}))) +                   
                               scale_y_continuous(expand = c(0, 0), 
                                                   limits = c(0, 8), 
                                                   breaks = seq(0, 8, 
                                                                by = 2)) +
                               theme_classic() +
                               theme(panel.grid.major = element_blank(),  
                                     panel.grid.minor = element_blank(), 
                                     axis.title =element_text(size=12),
                                     axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                                     axis.text=element_text(size=10),
                                     plot.title = element_text(size=12),
                                     legend.position="none")



# Heated  only 

MeanSE_TAOC_heated <-  Plate1_samples_CRE.final %>% 
                          dplyr::filter(!(Treatment %in% 'Ambient')) %>%
                          dplyr::mutate(Temperature = case_when(
                            Temperature.Number == 1 ~ "22",
                            Temperature.Number == 2 ~ "26",
                            Temperature.Number == 3 ~ "30",
                            Temperature.Number == 4 ~ "34",
                            Temperature.Number == 5 ~ "38",
                          )) %>%                               
                               ggplot(aes(x = factor(Temperature), 
                                   y = CRE_per_ugprotein, 
                                   colour=Treatment,
                                   group=Treatment)) +
                               scale_colour_manual(breaks=("Heated"), 
                                                            values=("orange")) +
                               geom_point(aes(colour = Treatment), 
                                              position = position_dodge2(width = 0.4)) + 
                               stat_summary(fun.y="mean", size = 0.8, color = "black",
                                            position = position_dodge2(width = 0.4)) +
                               stat_summary(fun.min = function(x) mean(x) - sd(x)/sqrt(length(x)), 
                                            fun.max = function(x) mean(x) + sd(x)/sqrt(length(x)),
                                            geom = 'errorbar', width = 0.25, size = 1,
                                            position = position_dodge2(width = 0.4)) +
                               labs(title="Total antioxidant capacity: Heated", 
                                    x ="Temperature",
                                    y=expression("TAOC"~(~uM~CRE[red]^{-1}*~ug~protein^{-1}))) +                   
                               scale_y_continuous(expand = c(0, 0), 
                                                   limits = c(0, 8), 
                                                   breaks = seq(0, 8, 
                                                                by = 2)) +
                               theme_classic() +
                               theme(panel.grid.major = element_blank(),  
                                     panel.grid.minor = element_blank(), 
                                     axis.title =element_text(size=12),
                                     axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
                                     axis.text=element_text(size=10),
                                     plot.title = element_text(size=12),
                                     legend.position="none")



# output 
pdf(paste("Output/Colorimetric_assays/Total_Antioxidant_Capacity/Total_Antoixidant_Capacity_AmbinetvHeated.pdf", sep =''), 
    width=10, 
    height=4)
ggarrange(MeanSE_TAOC_ambient, MeanSE_TAOC_heated)
dev.off()


pdf(paste("Output/Colorimetric_assays/Total_Antioxidant_Capacity/Total_Antioxidant_Capacity_Heated.pdf", sep =''), 
    width=5, 
    height=4)
print(MeanSE_TAOC_heated)
dev.off()

pdf(paste("Output/Colorimetric_assays/Total_Antioxidant_Capacity/Total_Antioxidant_Capacity_Ambient.pdf", sep =''), 
    width=5, 
    height=4)
print(MeanSE_TAOC_ambient)
dev.off()

```

